---
title: "Weekly Incidence: Allocation Pathways and Device Listings"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
date: "`r Sys.Date()`"
---

```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)


study_start <- as.Date("2024-09-01")   # first week begins
study_end   <- as.Date("2026-02-28")   # last full week before the 03/01 extract edge
policy_date <- as.Date("2025-09-01")
jama_pal <- c("#374E55", "#DF8F44", "#00A1D5", "#B24745", "#79AF97")

```
---

# Load data

```{r load-data}
JustFormHRStat1    <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat1.sas7bdat")
JustFormHRStat2    <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat2.sas7bdat")
JustFormHRStat3    <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat3.sas7bdat")
JustFormHRStat4    <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat4.sas7bdat")
JustFormHR         <- read_sas("/Users/chrisramos/Downloads/JustFormHR.sas7bdat")
StatHist_Thor      <- read_sas("/Users/chrisramos/Downloads/stathist_thor.sas7bdat")
JustFormHRDataLink <- read_sas("/Users/chrisramos/Downloads/JustFormHRDataLink.sas7bdat")
RiskStratDataHR    <- read_sas("/Users/chrisramos/Downloads/RiskStratDataHR.sas7bdat")
ThorSupportDevice  <- read_sas("/Users/chrisramos/Downloads/ThorSupportDevice.sas7bdat")
cand               <- read_sas("/Users/chrisramos/Downloads/cand_thor.sas7bdat")
```

---

# Weekly bins

Weeks are **inclusive 7-day blocks**: 2024-09-22 through 2024-09-28 is week 1,
2024-09-29 through 2024-10-05 is week 2, and so on. A date is assigned to the week
whose start is the most recent Sunday on or before it. Same logic for each month,
with each bin starting on the 1st of the month and ending on the day before the next
first of month.

```{r weekly-bins and monthly-bins}
week_starts <- seq.Date(from = study_start, to = study_end - 6, by = "week")

# map any date to the start of the week containing it; NA if outside the window
week_start <- function(d) {
  d <- as.Date(d)
  offset <- as.numeric(d - study_start)
  ws <- study_start + 7 * (offset %/% 7)
  ws[is.na(d) | d < study_start | d > study_end] <- NA
  ws
}

# sanity: boundaries land where we expect
tibble(
  date = as.Date(c("2024-09-22", "2024-09-28", "2024-09-29"))
) %>%
  mutate(assigned_week = week_start(date))

length(week_starts)
range(week_starts)

# ---- Monthly bins (parallel to the weekly ones) ----
month_starts <- seq.Date(
  from = as.Date(format(study_start, "%Y-%m-01")),
  to   = as.Date(format(study_end,   "%Y-%m-01")),
  by   = "month"
)
 
# assign any date to the first of its month; NA if outside the window
month_start <- function(d) {
  d  <- as.Date(d)
  ms <- as.Date(format(d, "%Y-%m-01"))
  ms[is.na(d) | d < study_start | d > study_end] <- NA
  ms
}
```

---

# Adult cohort and patient linkage

Unchanged: Daniel's adult rule, and the `JustId -> PX_ID` bridge.

```{r cohort-and-linkage}
adult_px <- cand %>%
  mutate(PX_ID = as.character(PX_ID)) %>%
  filter(CAN_AGE_IN_MONTHS_AT_LISTING / 12 > 17) %>%
  distinct(PX_ID)

justid_px_lookup_clean <- JustFormHRDataLink %>%
  mutate(JustId = as.character(JustId), WlregAuditId = as.character(WlregAuditId)) %>%
  select(JustId, WlregAuditId) %>%
  left_join(
    RiskStratDataHR %>%
      mutate(WlregAuditId = as.character(WlregAuditId)) %>%
      select(WlregAuditId, px_id),
    by = "WlregAuditId"
  ) %>%
  rename(PX_ID = px_id) %>%
  mutate(PX_ID = as.character(PX_ID)) %>%
  select(JustId, PX_ID) %>%
  distinct(JustId, .keep_all = TRUE)
```

---

# Device linkage

```{r device-decode}
device_type_decode <- tribble(
  ~DeviceTypeId, ~device_label,
  4L,            "TAH",
  8L,            "IABP",
  9L,            "VA ECMO",
  11L,           "Percutaneous Device",
  12L,           "Dischargeable VAD",
  13L,           "Non-Dischargeable VAD"
)

# The two device types this analysis tracks. "PEVAD" == percutaneous endovascular
# VAD == Impella, which the device file codes as "Percutaneous Device" (11) and the
# JustForms code as CriteriaMcsdEndovasSupp.
iabp_pevad <- c("IABP", "Percutaneous Device")
```

`device_long` links each device to a patient. `ThorSupportDevice` has no patient key;
the only bridge is `DeviceId -> JustFormHR$ThorSupportDevicePrimary -> JustId -> PX_ID`.

```{r device-long}
ThorSupportDevice <- ThorSupportDevice %>%
  mutate(
    DeviceId     = as.character(DeviceId),
    DeviceTypeId = as.integer(DeviceTypeId),
    ImplantDt    = as.Date(ImplantDt)
  )

device_justid_bridge <- JustFormHR %>%
  mutate(
    JustId                   = as.character(JustId),
    ThorSupportDevicePrimary = as.character(ThorSupportDevicePrimary)
  ) %>%
  filter(!is.na(ThorSupportDevicePrimary)) %>%
  select(DeviceId = ThorSupportDevicePrimary, JustId) %>%
  distinct()

device_long <- ThorSupportDevice %>%
  filter(!is.na(ImplantDt)) %>%
  inner_join(device_justid_bridge, by = "DeviceId") %>%
  left_join(justid_px_lookup_clean, by = "JustId") %>%
  left_join(device_type_decode, by = "DeviceTypeId") %>%
  filter(!is.na(PX_ID), !is.na(device_label)) %>%
  semi_join(adult_px, by = "PX_ID") %>%
  distinct(PX_ID, DeviceId, device_label, ImplantDt)
```

*Everything is listed based on ImplantDt <= CAN_LISTING_DT from now on; no more assumed explant dates*

---

# JustForm criteria

Status 2 (IABP, PEVAD, exception) comes from the justification forms. Status 3 high-dose inotrope is
rebuilt from `RiskStratDataHR` per Daniel's code. The only addition is attaching `JustId` to the 
risk-strat rows first, since that file carries `WlregAuditId` and `PX_ID` but not `JustId`.

```{r justform-shared}
JustFormHR <- JustFormHR %>%
  mutate(JustId = as.character(JustId), FormEffectiveDt = as.Date(FormEffectiveDt))

form_effective_lookup <- JustFormHR %>%
  select(JustId, FormEffectiveDt) %>%
  distinct(JustId, .keep_all = TRUE)

justform_core <- JustFormHR %>%
  transmute(JustId, exception = Exception %in% 1) %>%
  distinct(JustId, .keep_all = TRUE)
```

```{r stat3-inotrope-riskstrat}
# Only addition to the provided code: attach JustId (file has WlregAuditId + PX_ID).
risk_strat_data_hr <- RiskStratDataHR %>%
  mutate(WlregAuditId = as.character(WlregAuditId)) %>%
  rename(PX_ID = px_id) %>%                       # file uses lowercase px_id
  mutate(PX_ID = as.character(PX_ID)) %>%
  left_join(
    JustFormHRDataLink %>%
      mutate(JustId = as.character(JustId), WlregAuditId = as.character(WlregAuditId)) %>%
      select(JustId, WlregAuditId) %>% distinct(),
    by = "WlregAuditId"
  )

stat3_flags_riskstrat <- risk_strat_data_hr %>%
  mutate(
    single_high_dose =
      (CurrTherDobutamine >= 7.5) |
      (CurrTherMilrinone >= 0.50) |
      (CurrTherEpinephrine >= 0.02),

    dual_combo =
      (CurrTherDobutamine >= 3 & CurrTherMilrinone >= 0.25) |
      (CurrTherDobutamine >= 3 & CurrTherEpinephrine >= 0.01) |
      (CurrTherDobutamine >= 3 & CurrTherDopamine >= 3) |
      (CurrTherMilrinone >= 0.25 & CurrTherEpinephrine >= 0.01) |
      (CurrTherMilrinone >= 0.25 & CurrTherDopamine >= 3) |
      (CurrTherEpinephrine >= 0.01 & CurrTherDopamine >= 3),

    meets_inotrope_dose_criteria = coalesce(single_high_dose, FALSE) | coalesce(dual_combo, FALSE),

    on_support_at_hemo = HemoObtainedOnSupport %in% c(
      "Device Support", "Inotrope Support", "Inotrope Support, Device Support"
    ),

    ci_threshold_met = case_when(
      on_support_at_hemo ~ HemoCardiacIndex < 2.2,
      !on_support_at_hemo ~ HemoCardiacIndex < 1.8,
      TRUE ~ NA
    ),

    meets_cardiogenic_shock =
      HemoSbp < 90 &
      HemoPcwp > 15 &
      coalesce(ci_threshold_met, FALSE),

    has_hemo_monitoring = !is.na(HemoDataObtained) & HemoDataObtained != "",

    stat3_standard_hd_inotrope =
      coalesce(has_hemo_monitoring, FALSE) &
      coalesce(meets_cardiogenic_shock, FALSE) &
      coalesce(meets_inotrope_dose_criteria, FALSE)
  ) %>%
  group_by(PX_ID, JustId) %>%
  summarise(
    stat3_standard_hd_inotrope = any(stat3_standard_hd_inotrope, na.rm = TRUE),
    .groups = "drop"
  )

stat3_flags_riskstrat %>% count(stat3_standard_hd_inotrope)
```

```{r justform-events}
#Tracking events from justification forms
JustFormHRStat2 <- JustFormHRStat2 %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(form_effective_lookup, by = "JustId")

# RequestedCandStatCd lives on JustFormHR (not the StatN tables); pull it by JustId
requested_stat_lookup <- JustFormHR %>%
  mutate(JustId = as.character(JustId)) %>%
  select(JustId, RequestedCandStatCd) %>%
  distinct(JustId, .keep_all = TRUE)

status2_events <- JustFormHRStat2 %>%
  left_join(requested_stat_lookup, by = "JustId") %>%
  transmute(
    JustId, FormEffectiveDt, status_form = "Adult Status 2",
    RequestedCandStatCd,                              # carry status code to match against CAN_INIT_STAT at listing
    iabp          = CriteriaIabpSupport %in% 1,
    pevad         = CriteriaMcsdEndovasSupp %in% 1,
    inotrope_high = FALSE
  )

# Status 3 high-dose
status3_events <- stat3_flags_riskstrat %>%
  filter(stat3_standard_hd_inotrope) %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(form_effective_lookup, by = "JustId") %>%
  left_join(requested_stat_lookup, by = "JustId") %>%
  transmute(
    JustId, FormEffectiveDt, status_form = "Adult Status 3",
    RequestedCandStatCd,
    iabp          = FALSE,
    pevad         = FALSE,
    inotrope_high = TRUE
  )

justform_events <- bind_rows(status2_events, status3_events) %>%
  left_join(justform_core, by = "JustId") %>%
  left_join(justid_px_lookup_clean, by = "JustId") %>%
  mutate(PX_ID = as.character(PX_ID)) %>%
  filter(!is.na(PX_ID), !is.na(FormEffectiveDt)) %>%
  semi_join(adult_px, by = "PX_ID") %>%
  mutate(week = week_start(FormEffectiveDt)) %>%
  filter(!is.na(week))

# high-dose Status 3 events landing in-window
justform_events %>% filter(status_form == "Adult Status 3", inotrope_high) %>% nrow()
nrow(justform_events)

justform_events
```

---

# Graph 1: Weekly counts by allocation pathway

Five lines. Four are **form events**, tallied by `FormEffectiveDt`:

- Status 2 IABP
- Status 2 PEVAD (Impella / percutaneous)
- Status 3 high-dose inotropes; linked to just_form, counted from riskstrat
- Status 2 Exceptions (total)

The fifth is a **device event**, tallied by `ImplantDt`: every IABP or percutaneous
device implanted that week, regardless of what any form says.

```{r graph1-data}

#Start from each patient's initial listing dates, look up whether they had an
#IABP/PEVAD on or before the listing date -> flag the event as yes/no for device 
listing_device_flag <- cand %>%
  mutate(PX_ID = as.character(PX_ID), CAN_LISTING_DT = as.Date(CAN_LISTING_DT)) %>%
  semi_join(adult_px, by = "PX_ID") %>%
  select(PX_ID, CAN_LISTING_DT) %>%
  filter(!is.na(CAN_LISTING_DT)) %>%
  left_join(
    device_long %>%
      filter(device_label %in% iabp_pevad) %>%
      select(PX_ID, ImplantDt),
    by = "PX_ID",
    relationship = "many-to-many"
  ) %>%
  mutate(device_at_listing = !is.na(ImplantDt) & ImplantDt <= CAN_LISTING_DT) %>%
  group_by(PX_ID, CAN_LISTING_DT) %>%
  summarise(device_at_listing = any(device_at_listing), .groups = "drop")

# One row per NEW LISTING, carrying the initial-listing form's flags + device flag.
# The initial-listing justification is identified by matching the form's requested
# status to the candidate's initial status AND the form's effective date to the
# listing date. No need for governing form, since we are grabbing only justification
# on the date of initial listing (also checks for matching status).
listing_base <- cand %>%
  mutate(PX_ID = as.character(PX_ID), CAN_LISTING_DT = as.Date(CAN_LISTING_DT)) %>%
  semi_join(adult_px, by = "PX_ID") %>%
  filter(!is.na(CAN_LISTING_DT)) %>%
  select(PX_ID, CAN_LISTING_DT, CAN_INIT_STAT) %>%
  inner_join(
    justform_events %>%
      select(PX_ID, FormEffectiveDt, RequestedCandStatCd,
             iabp, pevad, inotrope_high, exception),
    by = "PX_ID",
    relationship = "many-to-many"
  ) %>%
  filter(
    CAN_INIT_STAT == RequestedCandStatCd,   # form's requested status == initial listing status
    FormEffectiveDt == CAN_LISTING_DT        # form effective on the listing date
  ) %>%
  group_by(PX_ID, CAN_LISTING_DT, CAN_INIT_STAT) %>%
  summarise(                                 # collapse in case a listing matched >1 form (should be rare/none)
    iabp          = any(iabp %in% TRUE),
    pevad         = any(pevad %in% TRUE),
    inotrope_high = any(inotrope_high %in% TRUE),
    exception     = any(exception %in% TRUE),
    .groups = "drop"
  ) %>%
  left_join(listing_device_flag, by = c("PX_ID", "CAN_LISTING_DT")) %>%
  mutate(
    device_at_listing = coalesce(device_at_listing, FALSE),
    week  = week_start(CAN_LISTING_DT),
    month = month_start(CAN_LISTING_DT)
  )

# did any listing match more than one form? (expect 0)
listing_base %>% count(PX_ID, CAN_LISTING_DT) %>% filter(n > 1)

# Name lines to prevent silent drops for no event cases
pathway_levels <- c(
  "Status 2: IABP",
  "Status 2: PEVAD",
  "Status 3: High-dose inotropes",
  "Status 2: Exception",
  "New listings with IABP/PEVAD at listing"
)

# Each line = count of new listings meeting one condition, per week
graph1_data <- bind_rows(
  listing_base %>% filter(CAN_INIT_STAT == 2120, iabp)          
  %>% count(week, name = "n") %>% mutate(pathway = "Status 2: IABP"),
  
  listing_base %>% filter(CAN_INIT_STAT == 2120, pevad)         
  %>% count(week, name = "n") %>% mutate(pathway = "Status 2: PEVAD"),
  
  listing_base %>% filter(CAN_INIT_STAT == 2130, inotrope_high) 
  %>% count(week, name = "n") %>% mutate(pathway = "Status 3: High-dose inotropes"),
  
  listing_base %>% filter(CAN_INIT_STAT == 2120, exception)     
  %>% count(week, name = "n") %>% mutate(pathway = "Status 2: Exception"),
  
  listing_base %>% filter(device_at_listing)                   
  %>% count(week, name = "n") %>% mutate(pathway = "New listings with IABP/PEVAD at listing")
) %>%
  mutate(pathway = factor(pathway, levels = pathway_levels)) %>%
  complete(week = week_starts, pathway = pathway_levels, fill = list(n = 0)) %>%
  arrange(pathway, week)

# all five present? (expect 5)
n_distinct(graph1_data$pathway)

graph1_data %>% group_by(pathway) %>% summarise(total_events = sum(n), .groups = "drop")
```

```{r graph1-plot, fig.width=12, fig.height=6}
ggplot(graph1_data, aes(x = week, y = n, color = pathway)) +
  geom_line() +
  geom_point(size = 0.9) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_color_manual(values = jama_pal) +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  labs(
    title = "Weekly Incidences",
    subtitle = "Form events at FormEffectiveDt; listings at CAN_LISTING_DT",
    x = "Week beginning", y = "Count of events", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  guides(color = guide_legend(nrow = 2))
```
```{r graph 1 monthly data} 
# Same lines as weekly, binned by month. listing_base already carries a `month` column.
graph1_data_monthly <- bind_rows(
  listing_base %>% filter(CAN_INIT_STAT == 2120, iabp)          
  %>% count(month, name = "n") %>% mutate(pathway = "Status 2: IABP"),
  
  listing_base %>% filter(CAN_INIT_STAT == 2120, pevad)         
  %>% count(month, name = "n") %>% mutate(pathway = "Status 2: PEVAD"),
  
  listing_base %>% filter(CAN_INIT_STAT == 2130, inotrope_high) 
  %>% count(month, name = "n") %>% mutate(pathway = "Status 3: High-dose inotropes"),
  
  listing_base %>% filter(CAN_INIT_STAT == 2120, exception)    
  %>% count(month, name = "n") %>% mutate(pathway = "Status 2: Exception"),
  
  listing_base %>% filter(device_at_listing)                   
  %>% count(month, name = "n") %>% mutate(pathway = "New listings with IABP/PEVAD at listing")
) %>%
  mutate(pathway = factor(pathway, levels = pathway_levels)) %>%
  complete(month = month_starts, pathway = pathway_levels, fill = list(n = 0)) %>%
  arrange(pathway, month)
 
graph1_data_monthly %>% group_by(pathway) %>% summarise(total_events = sum(n), .groups = "drop")
```

```{r graph1 monthly plots}
ggplot(graph1_data_monthly, aes(x = month, y = n, color = pathway)) +
  geom_line() +
  geom_point(size = 0.9) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_color_manual(values = jama_pal) +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  labs(
    title = "Monthly Incidences",
    subtitle = "Form events at FormEffectiveDt; listings at CAN_LISTING_DT",
    x = "Month beginning", y = "Count of events", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  guides(color = guide_legend(nrow = 2))
```

---

# Graph 2: Status 2 exception listings, with vs without IABP/PEVAD

The exact Status 2 exception population from the Graph 1 line, split by the device
flag. It is `listing_base` filtered to `CAN_INIT_STAT == 2120 & exception`, divided
by `device_at_listing` (IABP/PEVAD implanted on or before `CAN_LISTING_DT`).

A listing counts as **WITH device** if its `device_at_listing` flag is TRUE,
**WITHOUT** if FALSE. One row per listing, binned by `CAN_LISTING_DT`.

```{r graph2-data}
# Graph 2 is the exact Status 2 exception population from graph1_data, split by the
# device flag already on listing_base. Same rows, divided by device_at_listing.
group_levels <- c("Exception WITH IABP/PEVAD", "Exception WITHOUT IABP/PEVAD")

graph2_data <- listing_base %>%
  filter(CAN_INIT_STAT == 2120, exception, !is.na(week)) %>%
  mutate(group = if_else(device_at_listing,
                         "Exception WITH IABP/PEVAD",
                         "Exception WITHOUT IABP/PEVAD")) %>%
  count(week, group, name = "n") %>%
  complete(week = week_starts, group = group_levels, fill = list(n = 0)) %>%
  arrange(group, week)

# totals and share with device
graph2_data %>%
  group_by(group) %>%
  summarise(total = sum(n), .groups = "drop") %>%
  mutate(pct = round(100 * total / sum(total), 1))
```

```{r graph2-check}
# the two groups must sum to the total S2 exception listings per week (expect 0 rows)
graph2_data %>%
  group_by(week) %>%
  summarise(split_total = sum(n), .groups = "drop") %>%
  left_join(
    listing_base %>%
      filter(CAN_INIT_STAT == 2120, exception) %>%
      filter(!is.na(week)) %>%
      count(week, name = "exception_total"),
    by = "week"
  ) %>%
  mutate(exception_total = coalesce(exception_total, 0L)) %>%
  filter(split_total != exception_total)
```

```{r graph2-plot, fig.width=12, fig.height=6}
ggplot(graph2_data, aes(x = week, y = n, color = group)) +
  geom_line() +
  geom_point(size = 0.9) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_color_manual(values = c(
    "Exception WITH IABP/PEVAD"    = "#B24745",
    "Exception WITHOUT IABP/PEVAD" = "#374E55"
  )) +
  labs(
    title = "Weekly Status 2 Exception Listings, by IABP/PEVAD at Listing",
    subtitle = "Device evaluated at CAN_LISTING_DT; governing form determines exception status",
    x = "Week beginning", y = "Count of exception listings", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```
```{r graph2 monthly data}
graph2_data_monthly <- listing_base %>%
  filter(CAN_INIT_STAT == 2120, exception, !is.na(month)) %>%
  mutate(group = if_else(device_at_listing,
                         "Exception WITH IABP/PEVAD",
                         "Exception WITHOUT IABP/PEVAD")) %>%
  count(month, group, name = "n") %>%
  complete(month = month_starts, group = group_levels, fill = list(n = 0)) %>%
  arrange(group, month)
 
graph2_data_monthly %>%
  group_by(group) %>%
  summarise(total = sum(n), .groups = "drop") %>%
  mutate(pct = round(100 * total / sum(total), 1))
```

```{r graph2 monthly plots}
ggplot(graph2_data_monthly, aes(x = month, y = n, color = group)) +
  geom_line() +
  geom_point(size = 0.9) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_color_manual(values = c(
    "Exception WITH IABP/PEVAD"    = "#B24745",
    "Exception WITHOUT IABP/PEVAD" = "#374E55"
  )) +
  labs(
    title = "Monthly Status 2 Exception Listings, by IABP/PEVAD at Listing",
    subtitle = "Device evaluated at CAN_LISTING_DT; governing form determines exception status",
    x = "Month beginning", y = "Count of exception listings", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```

---

# Date coincidence check

Checking how often the listing date and justification form date does not line up, since those listed patients are getting dropped out of the analysis fully -> 3.2% of status 2/3 listings are dropped

```{r date-coincidence-check}
no_exact_prior <- cand %>%
  mutate(PX_ID = as.character(PX_ID), CAN_LISTING_DT = as.Date(CAN_LISTING_DT)) %>%
  semi_join(adult_px, by = "PX_ID") %>%
  filter(CAN_INIT_STAT %in% c(2120, 2130), !is.na(CAN_LISTING_DT)) %>%
  distinct(PX_ID, CAN_LISTING_DT, CAN_INIT_STAT) %>%
  left_join(
    justform_events %>% select(PX_ID, FormEffectiveDt, RequestedCandStatCd),
    by = "PX_ID", relationship = "many-to-many"
  ) %>%
  filter(CAN_INIT_STAT == RequestedCandStatCd) %>%
  group_by(PX_ID, CAN_LISTING_DT, CAN_INIT_STAT) %>%
  summarise(
    has_exact    = any(FormEffectiveDt == CAN_LISTING_DT),
    prior_form_dt = {
      before <- FormEffectiveDt[FormEffectiveDt < first(CAN_LISTING_DT)]
      if (length(before) == 0) as.Date(NA) else max(before)
    },
    .groups = "drop"
  ) %>%
  filter(!has_exact) %>%
  mutate(days_before_listing = as.numeric(CAN_LISTING_DT - prior_form_dt)) %>%
  arrange(days_before_listing) %>%
  select(PX_ID, CAN_INIT_STAT, CAN_LISTING_DT, prior_form_dt, days_before_listing)

no_exact_prior
#justification form is never preceding listing date; the issue is not a delay in
#registering patient as listed; have to investigate delay on registering justification
#form following listing

gap_dist <- cand %>%
  mutate(PX_ID = as.character(PX_ID), CAN_LISTING_DT = as.Date(CAN_LISTING_DT)) %>%
  semi_join(adult_px, by = "PX_ID") %>%
  filter(CAN_INIT_STAT %in% c(2120, 2130), !is.na(CAN_LISTING_DT)) %>%
  distinct(PX_ID, CAN_LISTING_DT, CAN_INIT_STAT) %>%
  left_join(
    justform_events %>% select(PX_ID, FormEffectiveDt, RequestedCandStatCd),
    by = "PX_ID", relationship = "many-to-many"
  ) %>%
  filter(CAN_INIT_STAT == RequestedCandStatCd) %>%
  group_by(PX_ID, CAN_LISTING_DT) %>%
  summarise(
    # signed gap of the NEAREST form to listing (min absolute distance, keeping sign)
    nearest_gap = {
      g <- as.numeric(FormEffectiveDt - first(CAN_LISTING_DT))
      g[which.min(abs(g))]
    },
    .groups = "drop"
  )

# counts per day-gap
gap_dist %>%
  filter(nearest_gap != 0, abs(nearest_gap) <= 30) %>%
  ggplot(aes(x = nearest_gap)) +
  geom_bar() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_x_continuous(breaks = seq(-30, 30, 5)) +
  labs(
    title = "Nearest justification form relative to listing date (exact matches excluded)",
    subtitle = "Negative = form before listing; positive = form after",
    x = "Days from listing date", y = "Number of listings"
  ) +
  theme_minimal(base_size = 12)

#a few patient's initial listing showed no justification form, and no form appears until
#mult. days later, likely to extend or change a status; excluding these people
```
