---
title: "JustForms"
output:
  html_document:
    self_contained: true
date: "2026-06-16"
---

```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)

# Snapshot window shortened to 2024-09-01 .. 2026-03-01. Starting in 2024-09 means every
# LVAD-supported candidate active in the window has a modern justification form, so the
# form-based numerators capture them directly -- no device backfill of pre-policy
# (pre-2018, auto-placed Status 4) patients is needed.
snapshot_start <- as.Date("2024-09-01")
snapshot_end   <- as.Date("2026-03-01")
policy_date    <- as.Date("2025-09-01")
```

## Load data

```{r load-data}
JustFormHRStat1 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat1.sas7bdat")
JustFormHRStat2 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat2.sas7bdat")
JustFormHRStat3 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat3.sas7bdat")
JustFormHRStat4 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat4.sas7bdat")
JustFormHR      <- read_sas("/Users/chrisramos/Downloads/JustFormHR.sas7bdat")
StatHist_Thor   <- read_sas("/Users/chrisramos/Downloads/stathist_thor.sas7bdat")
JustFormHRDataLink <- read_sas("/Users/chrisramos/Downloads/JustFormHRDataLink.sas7bdat")
RiskStratDataHR <- read_sas("/Users/chrisramos/Downloads/RiskStratDataHR.sas7bdat")
```

##Linking Just_ID to PX_ID

```{r}
justid_px_lookup <- JustFormHRDataLink %>%    #this dataset links JustID and WlregAuditID
  mutate(
    JustId = as.character(JustId),
    WlregAuditId = as.character(WlregAuditId)
  ) %>%

  select(JustId, WlregAuditId) %>%
  left_join(
    RiskStratDataHR %>%                      #this dataset link WlregAuditID to PX_ID
      mutate(
        WlregAuditId = as.character(WlregAuditId)
      ) %>%
      select(WlregAuditId, px_id),
    by = "WlregAuditId"                      #join our two simplified tables by common waitlist reg ID
  ) %>%
  rename(PX_ID = px_id)

#justid_px_lookup now has all 3 IDs joined by common Wlreg ID; If several JustIds share one WlregAuditId, they all get the same PX_ID

#sanity checks

#shows each just_id has a specific wlregaudit_id accounted for by a each row; as expected some patients have mult. of these ids
justid_px_lookup %>%
  summarise(
    n_rows = n(),
    n_JustId = n_distinct(JustId),
    n_WlregAuditId = n_distinct(WlregAuditId),
    n_PX_ID = n_distinct(PX_ID),
    missing_px_id = sum(is.na(PX_ID))
  )

#there are no duplicated just_ids
justid_px_lookup %>%
  count(JustId) %>%
  filter(n > 1)

#checking how many just_ids each px_id has associated with them
justid_px_lookup %>%
  group_by(PX_ID) %>%
  summarise(
    n_JustId = n_distinct(JustId),
    n_WlregAuditId = n_distinct(WlregAuditId),
    .groups = "drop"
  ) %>%
  arrange(desc(n_JustId))
```


## Create monthly snapshots

```{r monthly-snapshots}
## Monthly snapshot dates now come from status history, not JustForms

month_checkpoints <- seq.Date(
  from = snapshot_start,
  to = snapshot_end,
  by = "month"
)

add_active_month_snapshots <- function(df) {
  df %>%
    mutate(
      CANHX_BEGIN_DT = as.Date(CANHX_BEGIN_DT),
      CANHX_END_DT = as.Date(CANHX_END_DT),
      CAN_REM_DT = as.Date(CAN_REM_DT),
      CANHX_STAT_CD = as.character(CANHX_STAT_CD),
      
      # If no status-history end date exists, use removal date.
      # If neither exists, assume the status persists through snapshot_end.
      CANHX_END_DT_adjusted = case_when(
        !is.na(CANHX_END_DT) ~ CANHX_END_DT,
        is.na(CANHX_END_DT) & !is.na(CAN_REM_DT) ~ CAN_REM_DT,
        is.na(CANHX_END_DT) & is.na(CAN_REM_DT) ~ snapshot_end
      )
    ) %>%
    filter(
      CANHX_BEGIN_DT <= snapshot_end,
      CANHX_END_DT_adjusted >= snapshot_start
    ) %>%
    rowwise() %>%
    mutate(
      MonthSnapshot = list(
        month_checkpoints[
          month_checkpoints >= CANHX_BEGIN_DT &
            month_checkpoints <= CANHX_END_DT_adjusted
        ]
      )
    ) %>%
    ungroup() %>%
    tidyr::unnest(MonthSnapshot)
}

## No longer creating MonthSnapshot for JustForms. Using FormEffectiveDt so later we can ask for most recent JustForm as of MonthSnapshot.
## FormEffectiveDt is the form's effective date and lives ONLY in JustFormHR, keyed by JustId;
## it does NOT exist in the StatN tables. Build one authoritative JustId -> FormEffectiveDt
## lookup from JustFormHR and join it onto each StatN table, so all form-dating keys off it.
JustFormHR <- JustFormHR %>%
  mutate(
    JustId = as.character(JustId),
    FormEffectiveDt = as.Date(FormEffectiveDt)
  )

form_effective_lookup <- JustFormHR %>%
  select(JustId, FormEffectiveDt) %>%
  distinct(JustId, .keep_all = TRUE)

JustFormHRStat1 <- JustFormHRStat1 %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(form_effective_lookup, by = "JustId")

JustFormHRStat2 <- JustFormHRStat2 %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(form_effective_lookup, by = "JustId")

JustFormHRStat3 <- JustFormHRStat3 %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(form_effective_lookup, by = "JustId")

JustFormHRStat4 <- JustFormHRStat4 %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(form_effective_lookup, by = "JustId")

#clean JustId -> PX_ID lookup
justid_px_lookup_clean <- justid_px_lookup %>%
  mutate(
    JustId = as.character(JustId),
    PX_ID = as.character(PX_ID)
  ) %>%
  select(JustId, PX_ID) %>%
  distinct(JustId, .keep_all = TRUE)

#clean JustForm core
justform_core <- JustFormHR %>%
  transmute(
    JustId,
    RequestedCandStat_descrip,
    exception = Exception == 1
  ) %>%
  distinct(JustId, .keep_all = TRUE)

#build long JustForm table; keeps all forms up to snapshot_end. The most-recent-form-as-of-MonthSnapshot logic (below) will pick the governing form per month, so we retain forms dated before snapshot_start too (a form from, say, early 2024 can still be the most recent one governing 2024-09).
justform_status_long <- bind_rows(
  
  JustFormHRStat1 %>%
    transmute(
      JustId,
      FormEffectiveDt,
      status_form = "Adult Status 1",
      ecmo = CriteriaEcmoSupport == 1,
      iabp = FALSE,
      endovascular_mcsd = FALSE,
      lvad = FALSE,
      lvad_disc = FALSE,        # Status 3 only; FALSE here so bind_rows aligns
      inotrope_high = FALSE,    # Status 3 only
      inotrope_low = FALSE      # Status 4 only
    ),
  
  JustFormHRStat2 %>%
    transmute(
      JustId,
      FormEffectiveDt,
      status_form = "Adult Status 2",
      ecmo = FALSE,
      iabp = CriteriaIabpSupport == 1,
      endovascular_mcsd = CriteriaMcsdEndovasSupp == 1,   # impella == endovascular MCSD (Status 2 only)
      lvad = FALSE,
      lvad_disc = FALSE,
      inotrope_high = FALSE,
      inotrope_low = FALSE
    ),
  
  JustFormHRStat3 %>%
    transmute(
      JustId,
      FormEffectiveDt,
      status_form = "Adult Status 3",
      ecmo = FALSE,
      iabp = CriteriaIabpSupport == 1,
      endovascular_mcsd = FALSE,
      lvad = FALSE,
      lvad_disc = CriteriaLvadDiscSupport == 1,            # LVAD discretionary (Status 3)
      inotrope_high = CriteriaInotropeSupport == 1,        # high-dose inotropes (Status 3 file)
      inotrope_low = FALSE
    ),
  
  JustFormHRStat4 %>%
    transmute(
      JustId,
      FormEffectiveDt,
      status_form = "Adult Status 4",
      ecmo = FALSE,
      iabp = FALSE,
      endovascular_mcsd = FALSE,
      lvad = CriteriaLvadSupport == 1,
      lvad_disc = FALSE,
      inotrope_high = FALSE,
      inotrope_low = CriteriaInotropeSupport == 1          # low-dose inotropes (Status 4 file)
    )
) %>%
  filter(
    !is.na(FormEffectiveDt),
    FormEffectiveDt <= snapshot_end
  ) %>%
  left_join(
    justid_px_lookup_clean,
    by = "JustId"
  ) %>%
  left_join(
    justform_core,
    by = "JustId"
  ) %>%
  mutate(
    PX_ID = as.character(PX_ID)
  )

```

##Prepare stathist_thor and get a total monthly candidate count across all 6 statuses
```{r}
## stathist_thor monthly active denominator

#heart-only status history rows, expanded to active first-of-month snapshots
stathist_filtered <- StatHist_Thor %>%
  mutate(
    PX_ID = as.character(PX_ID)
  ) %>%
  filter(WL_ORG == "HR") %>%
  add_active_month_snapshots()

#keep adult statuses 1-6 and label them
stathist_monthly <- stathist_filtered %>%
  filter(CANHX_STAT_CD %in% c(
    "2110", "2120", "2130", "2140", "2150", "2160"
  )) %>%
  mutate(
    CANHX_STAT_label = recode(
      CANHX_STAT_CD,
      "2110" = "Adult Status 1",
      "2120" = "Adult Status 2",
      "2130" = "Adult Status 3",
      "2140" = "Adult Status 4",
      "2150" = "Adult Status 5",
      "2160" = "Adult Status 6"
    )
  )

#gets patient/status/month frame from stathist
active_status_patient_months <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  distinct(
    PX_ID,
    MonthSnapshot,
    status_form = CANHX_STAT_label
  )

#for each active patient/status/month, take the most recent JustForm
justform_patient_month <- active_status_patient_months %>%
  inner_join(
    justform_status_long %>%
      filter(
        !is.na(PX_ID),
        !is.na(FormEffectiveDt)
      ) %>%
      select(
        PX_ID,
        JustId,
        FormEffectiveDt,
        status_form,
        RequestedCandStat_descrip,
        ecmo,
        iabp,
        endovascular_mcsd,
        lvad,
        lvad_disc,
        inotrope_high,
        inotrope_low,
        exception
      ),
    by = c("PX_ID", "status_form"),
    relationship = "many-to-many"
  ) %>%
  filter(
    FormEffectiveDt <= MonthSnapshot
  ) %>%
  group_by(PX_ID, MonthSnapshot, status_form) %>%
  arrange(
    desc(FormEffectiveDt),
    desc(suppressWarnings(as.numeric(JustId))),
    .by_group = TRUE
  ) %>%
  slice(1) %>%
  ungroup()

#check that there is only one JustForm row per patient/status/month.
justform_patient_month %>%
  count(PX_ID, MonthSnapshot, status_form) %>%
  filter(n > 1)
```


## Total monthly denominator (total candidates listed per month (stat 1-6))

```{r total-denominator}
#Total monthly denominator: unique active Status 1-6 heart candidates
all_patient_monthly_counts <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  group_by(MonthSnapshot) %>%
  summarise(
    n_all = n_distinct(PX_ID),
    .groups = "drop"
  )
```

## Monthly numerators and proportions (how much of each device)

```{r monthly-numerators}
#how many (individuals, month) have criterion of interest (adds new label, keeps status value as well)
summarise_patient_criterion <- function(status_value, flag_col, support_label) {
  justform_patient_month %>%
    group_by(MonthSnapshot) %>%
    summarise(
      support_type = support_label,
      n_yes = sum(
        status_form == status_value & .data[[flag_col]],
        na.rm = TRUE
      ),
      .groups = "drop"
    )
}

#use function to generate all numerators desired + label
criteria_numerators <- bind_rows(
  summarise_patient_criterion("Adult Status 1", "ecmo", "Status 1: ECMO"),
  summarise_patient_criterion("Adult Status 2", "iabp", "Status 2: IABP"),
  summarise_patient_criterion("Adult Status 2", "endovascular_mcsd", "Status 2: Endovascular MCSD"),
  summarise_patient_criterion("Adult Status 3", "iabp", "Status 3: IABP"),
  summarise_patient_criterion("Adult Status 4", "lvad", "Status 4: LVAD")
)

support_levels <- c(
  "Status 1: ECMO",
  "Status 2: IABP",
  "Status 2: Endovascular MCSD",
  "Status 3: IABP",
  "Status 4: LVAD"
)

#generate numerators for each criteria
criteria_month_props <- criteria_numerators %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = all_patient_monthly_counts$MonthSnapshot,
    support_type = support_levels,
    fill = list(n_yes = 0)
  ) %>%
  left_join(all_patient_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_all
  )
```

## Overall support-type proportions

```{r overall-support-proportions}
overall_support_month_props <- criteria_numerators %>%
  mutate(
    support_type = case_when(
      support_type == "Status 1: ECMO" ~ "ECMO",
      support_type %in% c("Status 2: IABP", "Status 3: IABP") ~ "IABP",
      support_type == "Status 2: Endovascular MCSD" ~ "Endovascular MCSD",
      support_type == "Status 4: LVAD" ~ "LVAD",
      TRUE ~ support_type
    )
  ) %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = all_patient_monthly_counts$MonthSnapshot,
    support_type = c("ECMO", "IABP", "Endovascular MCSD", "LVAD"),
    fill = list(n_yes = 0)
  ) %>%
  left_join(all_patient_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_all
  )
```

```{r overall-support-plot}
ggplot(
  overall_support_month_props,
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.50),
    breaks = seq(0, 0.50, 0.05),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Active Heart Candidates by Support Type",
    x = "Month Snapshot",
    y = "Proportion of Active Status 1–6 Candidates",
    color = "Support Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Same graphs w/o lvad on there

```{r}
ggplot(
  overall_support_month_props %>%
    filter(support_type != "LVAD"),
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.10),
    breaks = seq(0, 0.10, 0.01),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Active Heart Candidates by Support Type",
    subtitle = "LVAD excluded",
    x = "Month Snapshot",
    y = "Proportion of Active Status 1–6 Candidates",
    color = "Support Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```


## Within-status numerator helper

```{r within-status-helper}
# Reusable within-status builders. Pattern used by ALL of Status 1-4 below:
#   numerator   = active-in-status patient-months whose governing JustForm has the flag
#   denominator = active-in-status patient-months from stathist
# Every status section uses the same three helpers and the same plot styling, so the
# graphs are directly comparable.
#
# Returns (for each helper): MonthSnapshot, support_type, n_yes, n_status, proportion.

# denominator for any status, from stathist
status_monthly_denominator <- function(status_label) {
  stathist_monthly %>%
    filter(CANHX_STAT_label == status_label, !is.na(PX_ID)) %>%
    distinct(PX_ID, MonthSnapshot) %>%
    group_by(MonthSnapshot) %>%
    summarise(n_status = n_distinct(PX_ID), .groups = "drop")
}

# one within-status criterion: distinct patients whose governing form has `flag_col`
within_status_criterion <- function(criteria_rows, denom, flag_col, support_label) {
  criteria_rows %>%
    group_by(MonthSnapshot) %>%
    summarise(
      n_yes = n_distinct(PX_ID[.data[[flag_col]]]),
      .groups = "drop"
    ) %>%
    complete(MonthSnapshot = denom$MonthSnapshot, fill = list(n_yes = 0)) %>%
    mutate(support_type = support_label) %>%
    left_join(denom, by = "MonthSnapshot") %>%
    mutate(proportion = n_yes / n_status) %>%
    select(MonthSnapshot, support_type, n_yes, n_status, proportion)
}

# within-status count over an arbitrary row-predicate (used for "Other" and for combined
# device = IABP or impella). `predicate_fn` is a function of the data frame returning a
# logical vector (e.g. function(d) d$iabp | d$endovascular_mcsd); distinct patients
# satisfying it are counted. Using a plain function avoids any tidy-eval/rlang fragility.
within_status_union <- function(criteria_rows, denom, predicate_fn, support_label) {
  df <- criteria_rows
  df$.match <- predicate_fn(df)        # logical vector over all rows, computed up front
  df %>%
    group_by(MonthSnapshot) %>%
    summarise(
      n_yes = n_distinct(PX_ID[.match]),
      .groups = "drop"
    ) %>%
    complete(MonthSnapshot = denom$MonthSnapshot, fill = list(n_yes = 0)) %>%
    mutate(support_type = support_label) %>%
    left_join(denom, by = "MonthSnapshot") %>%
    mutate(proportion = n_yes / n_status) %>%
    select(MonthSnapshot, support_type, n_yes, n_status, proportion)
}

# shared plot styling so every within-status graph matches
within_status_plot <- function(df, plot_title, y_label = "Proportion of Active Candidates",
                                y_max = 1.0, y_step = 0.1) {
  ggplot(df, aes(x = MonthSnapshot, y = proportion, color = support_type)) +
    geom_line() +
    geom_point(size = 1) +
    geom_vline(xintercept = policy_date, linetype = "dashed") +
    scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
    scale_y_continuous(
      limits = c(0, y_max),
      breaks = seq(0, y_max, y_step),
      labels = scales::percent_format()
    ) +
    labs(title = plot_title, x = "Month Snapshot", y = y_label, color = "Criterion") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


## Status 4: LVAD vs low-dose inotropes (within Status 4)

```{r status4-criteria-proportions}
status4_denom <- status_monthly_denominator("Adult Status 4")

status4_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 4")

status4_props <- bind_rows(
  within_status_criterion(status4_criteria_rows, status4_denom, "lvad",         "Status 4: LVAD"),
  within_status_criterion(status4_criteria_rows, status4_denom, "inotrope_low", "Status 4: Low-dose inotropes")
)

status4_props
```

```{r status4-criteria-plot}
within_status_plot(
  status4_props,
  "Monthly Proportion of Status 4 Candidates by Criterion",
  y_label = "Proportion of Active Status 4 Candidates"
)
```


## Status 3: LVAD discretionary vs high-dose inotropes (within Status 3)

LVAD complications criterion is deferred for now (to be added later).

```{r status3-criteria-proportions}
status3_denom <- status_monthly_denominator("Adult Status 3")

status3_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 3")

status3_props <- bind_rows(
  within_status_criterion(status3_criteria_rows, status3_denom, "lvad_disc",     "Status 3: LVAD discretionary"),
  within_status_criterion(status3_criteria_rows, status3_denom, "inotrope_high", "Status 3: High-dose inotropes")
  # TODO: add "Status 3: LVAD complications" once that criterion source is identified
)

status3_props
```

```{r status3-criteria-plot}
within_status_plot(
  status3_props,
  "Monthly Proportion of Status 3 Candidates by Criterion",
  y_label = "Proportion of Active Status 3 Candidates"
)

View(JustFormHRStat3)
```


## Status 2: IABP, impella, exceptions, and Other (within Status 2)

Two plots, same format: (A) IABP / impella(=Endovascular MCSD) / Exception / Other
broken out separately; (B) the combined device (IABP or impella) / Exception / Other.
"Other" = active Status 2 patients whose governing form is none of {IABP, impella,
exception}. The overlap check below confirms there is no overlap between the device
criteria and exceptions (and re-confirms IABP and impella do not overlap), so Other is a
clean complement.

```{r status2-overlap-check}
status2_denom <- status_monthly_denominator("Adult Status 2")

status2_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 2")

status2_overlap_check <- status2_criteria_rows %>%
  summarise(
    n_rows                 = n(),
    n_iabp                 = sum(iabp, na.rm = TRUE),
    n_impella_endovas      = sum(endovascular_mcsd, na.rm = TRUE),
    n_exception            = sum(exception, na.rm = TRUE),
    n_device               = sum((iabp | endovascular_mcsd), na.rm = TRUE),
    n_device_and_exception = sum((iabp | endovascular_mcsd) & exception, na.rm = TRUE),
    n_iabp_and_impella     = sum(iabp & endovascular_mcsd, na.rm = TRUE)
  )

status2_overlap_check
```

```{r status2-broken-out-proportions}
# (A) IABP / impella / exception / other, broken out
status2_props_brokenout <- bind_rows(
  within_status_criterion(status2_criteria_rows, status2_denom, "iabp",              "Status 2: IABP"),
  within_status_criterion(status2_criteria_rows, status2_denom, "endovascular_mcsd", "Status 2: Impella (Endovascular MCSD)"),
  within_status_criterion(status2_criteria_rows, status2_denom, "exception",         "Status 2: Exception"),
  within_status_union(status2_criteria_rows, status2_denom,
                      function(d) !d$exception & !d$iabp & !d$endovascular_mcsd, "Status 2: Other")
)

status2_props_brokenout
```

```{r status2-broken-out-plot}
within_status_plot(
  status2_props_brokenout,
  "Monthly Proportion of Status 2 Candidates by Criterion (broken out)",
  y_label = "Proportion of Active Status 2 Candidates"
)
```

```{r status2-combined-proportions}
# (B) combined device (IABP or impella) / exception / other
status2_props_combined <- bind_rows(
  within_status_union(status2_criteria_rows, status2_denom,
                      function(d) d$iabp | d$endovascular_mcsd, "Status 2: IABP or Impella"),
  within_status_criterion(status2_criteria_rows, status2_denom, "exception", "Status 2: Exception"),
  within_status_union(status2_criteria_rows, status2_denom,
                      function(d) !d$exception & !d$iabp & !d$endovascular_mcsd, "Status 2: Other")
)

status2_props_combined
```

```{r status2-combined-plot}
within_status_plot(
  status2_props_combined,
  "Monthly Proportion of Status 2 Candidates: Device vs Exception vs Other",
  y_label = "Proportion of Active Status 2 Candidates"
)
```


## Status 1: ECMO, exceptions, and Other (within Status 1)

"Other" = active Status 1 patients whose governing form is neither ECMO nor exception.
The overlap check below confirms ECMO and exception do not overlap, so Other is a clean
complement.

```{r status1-overlap-check}
status1_denom <- status_monthly_denominator("Adult Status 1")

status1_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 1")

status1_overlap_check <- status1_criteria_rows %>%
  summarise(
    n_rows               = n(),
    n_ecmo               = sum(ecmo, na.rm = TRUE),
    n_exception          = sum(exception, na.rm = TRUE),
    n_ecmo_and_exception = sum(ecmo & exception, na.rm = TRUE),
    n_ecmo_or_exception  = sum(ecmo | exception, na.rm = TRUE)
  )

status1_overlap_check
```

```{r status1-proportions}
status1_props <- bind_rows(
  within_status_criterion(status1_criteria_rows, status1_denom, "ecmo",      "Status 1: ECMO"),
  within_status_criterion(status1_criteria_rows, status1_denom, "exception", "Status 1: Exception"),
  within_status_union(status1_criteria_rows, status1_denom,
                      function(d) !d$ecmo & !d$exception, "Status 1: Other")
)

status1_props
```

```{r status1-plot}
within_status_plot(
  status1_props,
  "Monthly Proportion of Status 1 Candidates by Criterion",
  y_label = "Proportion of Active Status 1 Candidates"
)
```

```{r}
# Per-month sum of the tracked proportions within Status 1 and Status 2.
# If the criteria within a status partition its denominator cleanly, these sums
# should be ~1.0 each month (Other is the complement, so it closes the gap).
status12_prop_sums <- bind_rows(
  status1_props %>%
    group_by(MonthSnapshot) %>%
    summarise(
      status        = "Status 1",
      sum_n_yes     = sum(n_yes),
      n_status      = first(n_status),
      sum_proportion = sum(proportion),
      .groups = "drop"
    ),
  status2_props_brokenout %>%
    group_by(MonthSnapshot) %>%
    summarise(
      status        = "Status 2",
      sum_n_yes     = sum(n_yes),
      n_status      = first(n_status),
      sum_proportion = sum(proportion),
      .groups = "drop"
    )
) %>%
  arrange(status, MonthSnapshot) %>%
  relocate(status, MonthSnapshot)

status12_prop_sums
```

