---
title: "JustForms"
output:
  html_document:
    self_contained: true
date: "2026-06-16"
---

```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)

snapshot_start <- as.Date("2024-09-01")
snapshot_end   <- as.Date("2026-03-01")
policy_date    <- as.Date("2025-09-01")
```

## Load data

```{r load-data}
JustFormHRStat1 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat1.sas7bdat")
JustFormHRStat2 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat2.sas7bdat")
JustFormHRStat3 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat3.sas7bdat")
JustFormHRStat4 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat4.sas7bdat")
JustFormHR      <- read_sas("/Users/chrisramos/Downloads/JustFormHR.sas7bdat")
StatHist_Thor   <- read_sas("/Users/chrisramos/Downloads/stathist_thor.sas7bdat")
JustFormHRDataLink <- read_sas("/Users/chrisramos/Downloads/JustFormHRDataLink.sas7bdat")
RiskStratDataHR <- read_sas("/Users/chrisramos/Downloads/RiskStratDataHR.sas7bdat")
ThorSupportDevice <- read_sas("/Users/chrisramos/Downloads/ThorSupportDevice.sas7bdat")

ColumnDesc <- read_sas("/Users/chrisramos/Downloads/ColumnDescriptions (1).sas7bdat")
```

##Linking Just_ID to PX_ID

```{r}
justid_px_lookup <- JustFormHRDataLink %>%    #this dataset links JustID and WlregAuditID -> justification form linked to waitlist registration
  mutate(
    JustId = as.character(JustId),
    WlregAuditId = as.character(WlregAuditId)
  ) %>%

#only interested in Just and Wlreg IDs in the DataLink table
  select(JustId, WlregAuditId) %>%
  left_join(
    RiskStratDataHR %>%                      #this dataset link WlregAuditID to PX_ID -> waitlist registrations linked to patient 
      mutate(
        WlregAuditId = as.character(WlregAuditId)
      ) %>%
      select(WlregAuditId, px_id),
    by = "WlregAuditId"                      #join our two simplified tables by common waitlist reg ID (we now have all waitlist IDs)
  ) %>%
  rename(PX_ID = px_id)

#justid_px_lookup now has all 3 IDs joined by common Wlreg ID; If several JustIds share one WlregAuditId, they all get the same PX_ID

# sanity checks

#shows each just_id has a specific wlregaudit_id accounted for by a each row; as expected some patients have mult. of these ids
justid_px_lookup %>%
  summarise(
    n_rows = n(),
    n_JustId = n_distinct(JustId),
    n_WlregAuditId = n_distinct(WlregAuditId),
    n_PX_ID = n_distinct(PX_ID),
    missing_px_id = sum(is.na(PX_ID))
  )

# there are no duplicated just_ids
justid_px_lookup %>%
  count(JustId) %>%
  filter(n > 1)

#checking how many just_ids each px_id has associated with them
justid_px_lookup %>%
  group_by(PX_ID) %>%
  summarise(
    n_JustId = n_distinct(JustId),
    n_WlregAuditId = n_distinct(WlregAuditId),
    .groups = "drop"
  ) %>%
  arrange(desc(n_JustId))
```


## Create monthly snapshots

```{r monthly-snapshots}
## Monthly snapshot dates now come from status history, not JustForms

month_checkpoints <- seq.Date(
  from = snapshot_start,
  to = snapshot_end,
  by = "month"
)

add_active_month_snapshots <- function(df) {
  df %>%
    mutate(
      CANHX_BEGIN_DT = as.Date(CANHX_BEGIN_DT),
      CANHX_END_DT = as.Date(CANHX_END_DT),
      CAN_REM_DT = as.Date(CAN_REM_DT),
      CANHX_STAT_CD = as.character(CANHX_STAT_CD),
      
      # If no status-history end date exists, use removal date.
      # If neither exists, assume the status persists through snapshot_end.
      CANHX_END_DT_adjusted = case_when(
        !is.na(CANHX_END_DT) ~ CANHX_END_DT,
        is.na(CANHX_END_DT) & !is.na(CAN_REM_DT) ~ CAN_REM_DT,
        is.na(CANHX_END_DT) & is.na(CAN_REM_DT) ~ snapshot_end
      )
    ) %>%
    filter(
      CANHX_BEGIN_DT <= snapshot_end,
      CANHX_END_DT_adjusted >= snapshot_start
    ) %>%
    rowwise() %>%
    mutate(
      MonthSnapshot = list(
        month_checkpoints[
          month_checkpoints >= CANHX_BEGIN_DT &
            month_checkpoints <= CANHX_END_DT_adjusted
        ]
      )
    ) %>%
    ungroup() %>%
    tidyr::unnest(MonthSnapshot)
}

## No longer creating MonthSnapshot for JustForms. Using ChangeDate so later we can ask for most recent JustForm as of MonthSnapshot
JustFormHRStat1 <- JustFormHRStat1 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat2 <- JustFormHRStat2 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat3 <- JustFormHRStat3 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat4 <- JustFormHRStat4 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHR <- JustFormHR %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

#convert JustForm dates/IDs only. Do NOT create MonthSnapshot for JustForms.
JustFormHRStat1 <- JustFormHRStat1 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat2 <- JustFormHRStat2 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat3 <- JustFormHRStat3 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat4 <- JustFormHRStat4 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHR <- JustFormHR %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

#clean JustId -> PX_ID lookup
justid_px_lookup_clean <- justid_px_lookup %>%
  mutate(
    JustId = as.character(JustId),
    PX_ID = as.character(PX_ID)
  ) %>%
  select(JustId, PX_ID) %>%
  distinct(JustId, .keep_all = TRUE)

#clean ThorSupportDevice (collapsing vad and percu brand into one column also)
ThorSupportDevice_clean <- ThorSupportDevice %>%
  mutate(
    DeviceId = trimws(as.character(DeviceId)),
    ImplantDt = as.Date(ImplantDt),
    ChangeDate_device = as.Date(ChangeDate),
    brand_id = coalesce(
      as.numeric(VadBrandId),
      as.numeric(PercuBrandId)
    ),
    device_found_in_thor = TRUE
  )

#bridging DevicePrimary (justformhr) -> DeviceID (thorsupportdevice)
justform_device_link <- JustFormHR %>%
  mutate(
    JustId = trimws(as.character(JustId)),
    ChangeDate_justform = as.Date(ChangeDate),
    DeviceId = trimws(as.character(ThorSupportDevicePrimary))
  ) %>%
  mutate(
    DeviceId = na_if(DeviceId, ""),
    DeviceId = na_if(DeviceId, "NA")
  ) %>%
  filter(!is.na(DeviceId)) %>%
  select(
    JustId,
    ChangeDate_justform,
    DeviceId
  ) %>%
  distinct()

#get px_id in through justid + add new columns from thorsupportdevice
thor_device_patient_link <- justform_device_link %>%
  left_join(
    justid_px_lookup_clean,
    by = "JustId"
  ) %>%
  left_join(
    ThorSupportDevice_clean %>%
      select(
        DeviceId,
        DeviceTypeId,
        VadBrandId,
        PercuBrandId,
        brand_id,
        VentricularSupportId,
        ImplantDt,
        ChangeDate_device,
        device_found_in_thor
      ),
    by = "DeviceId"
  ) %>%
  mutate(
    device_found_in_thor = replace_na(device_found_in_thor, FALSE),
    PX_ID = as.character(PX_ID)
  )

durable_lvad_brand_ids <- c(202, 205, 206, 207, 208, 209, 210, 212, 213, 214, 223, 224, 233, 236, 239, 240, 312, 313, 314, 315, 316, 319, 322, 327, 330, 333, 334)

thor_device_patient_link <- thor_device_patient_link %>%
  mutate(
    brand_id = as.numeric(brand_id),
    durable_lvad = if_else(
      brand_id %in% durable_lvad_brand_ids,
      1L,
      0L,
      missing = 0L
    )
  )

thor_device_patient_link %>%
  count(brand_id, durable_lvad) %>%
  arrange(brand_id)

#clean JustForm core
justform_core <- JustFormHR %>%
  transmute(
    JustId,
    RequestedCandStat_descrip,
    exception = Exception == 1
  ) %>%
  distinct(JustId, .keep_all = TRUE)

#build long JustForm table; this keeps all forms up to snapshot_end, including pre-2019 forms (important for LVAD counts).
justform_status_long <- bind_rows(
  
  JustFormHRStat1 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 1",
      ecmo = CriteriaEcmoSupport == 1,
      iabp = FALSE,
      endovascular_mcsd = FALSE,
      lvad = FALSE
    ),
  
  JustFormHRStat2 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 2",
      ecmo = FALSE,
      iabp = CriteriaIabpSupport == 1,
      endovascular_mcsd = CriteriaMcsdEndovasSupp == 1,
      lvad = FALSE
    ),
  
  JustFormHRStat3 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 3",
      ecmo = FALSE,
      iabp = CriteriaIabpSupport == 1,
      endovascular_mcsd = FALSE,
      lvad = FALSE
    ),
  
  JustFormHRStat4 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 4",
      ecmo = FALSE,
      iabp = FALSE,
      endovascular_mcsd = FALSE,
      lvad = CriteriaLvadSupport == 1
    )
) %>%
  filter(
    !is.na(ChangeDate),
    ChangeDate <= snapshot_end
  ) %>%
  left_join(
    justid_px_lookup_clean,
    by = "JustId"
  ) %>%
  left_join(
    justform_core,
    by = "JustId"
  ) %>%
  mutate(
    PX_ID = as.character(PX_ID)
  )

#checking device id bridge
thor_device_patient_link %>%
  summarise(
    n_rows = n(),
    n_JustId = n_distinct(JustId),
    n_DeviceId = n_distinct(DeviceId),
    n_PX_ID = n_distinct(PX_ID),
    n_DeviceId_found_in_ThorSupportDevice = n_distinct(DeviceId[device_found_in_thor]),
    prop_DeviceId_found = n_DeviceId_found_in_ThorSupportDevice / n_DeviceId,
    missing_PX_ID = sum(is.na(PX_ID)),
    missing_ImplantDt = sum(is.na(ImplantDt)),
    missing_brand_id = sum(is.na(brand_id))
  )

thor_device_patient_link %>%
  group_by(DeviceId) %>%
  summarise(
    n_PX_ID = n_distinct(PX_ID),
    .groups = "drop"
  ) %>%
  filter(n_PX_ID > 1)

```

##Prepare stathist_thor and get a total monthly candidate count across all 6 statuses
```{r}
## stathist_thor monthly active denominator

#heart-only status history rows, expanded to active first-of-month snapshots
stathist_filtered <- StatHist_Thor %>%
  mutate(
    PX_ID = as.character(PX_ID)
  ) %>%
  filter(WL_ORG == "HR") %>%
  add_active_month_snapshots()

#keep adult statuses 1-6 and label them
stathist_monthly <- stathist_filtered %>%
  filter(CANHX_STAT_CD %in% c(
    "2110", "2120", "2130", "2140", "2150", "2160"
  )) %>%
  mutate(
    CANHX_STAT_label = recode(
      CANHX_STAT_CD,
      "2110" = "Adult Status 1",
      "2120" = "Adult Status 2",
      "2130" = "Adult Status 3",
      "2140" = "Adult Status 4",
      "2150" = "Adult Status 5",
      "2160" = "Adult Status 6"
    )
  )

#gets patient/status/month frame from stathist
active_status_patient_months <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  distinct(
    PX_ID,
    MonthSnapshot,
    status_form = CANHX_STAT_label
  )

#for each active patient/status/month, take the most recent JustForm
justform_patient_month <- active_status_patient_months %>%
  inner_join(
    justform_status_long %>%
      filter(
        !is.na(PX_ID),
        !is.na(ChangeDate)
      ) %>%
      select(
        PX_ID,
        JustId,
        ChangeDate,
        status_form,
        RequestedCandStat_descrip,
        ecmo,
        iabp,
        endovascular_mcsd,
        lvad,
        exception
      ),
    by = c("PX_ID", "status_form"),
    relationship = "many-to-many"
  ) %>%
  filter(
    ChangeDate <= MonthSnapshot
  ) %>%
  group_by(PX_ID, MonthSnapshot, status_form) %>%
  arrange(
    desc(ChangeDate),
    desc(suppressWarnings(as.numeric(JustId))),
    .by_group = TRUE
  ) %>%
  slice(1) %>%
  ungroup()

#check that there is only one JustForm row per patient/status/month.
justform_patient_month %>%
  count(PX_ID, MonthSnapshot, status_form) %>%
  filter(n > 1)

#Total monthly denominator: unique active Status 1-6 heart candidates
all_patient_monthly_counts <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  group_by(MonthSnapshot) %>%
  summarise(
    n_all = n_distinct(PX_ID),
    .groups = "drop"
  )
```


## Total monthly denominator (total candidates listed per month (stat 1-6))

```{r total-denominator}
all_patient_monthly_counts <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  group_by(MonthSnapshot) %>%
  summarise(
    n_all = n_distinct(PX_ID),
    .groups = "drop"
  )
```

## Monthly numerators and proportions (how much of each device)

```{r monthly-numerators}
#how many (individuals, month) have criterion of interest (adds new label, keeps status value as well)
summarise_patient_criterion <- function(status_value, flag_col, support_label) {
  justform_patient_month %>%
    group_by(MonthSnapshot) %>%
    summarise(
      support_type = support_label,
      n_yes = sum(
        status_form == status_value & .data[[flag_col]],
        na.rm = TRUE
      ),
      .groups = "drop"
    )
}

#use function to generate all numerators desired + label
criteria_numerators <- bind_rows(
  summarise_patient_criterion("Adult Status 1", "ecmo", "Status 1: ECMO"),
  summarise_patient_criterion("Adult Status 2", "iabp", "Status 2: IABP"),
  summarise_patient_criterion("Adult Status 2", "endovascular_mcsd", "Status 2: Endovascular MCSD"),
  summarise_patient_criterion("Adult Status 3", "iabp", "Status 3: IABP"),
  summarise_patient_criterion("Adult Status 4", "lvad", "Status 4: LVAD")
)

support_levels <- c(
  "Status 1: ECMO",
  "Status 2: IABP",
  "Status 2: Endovascular MCSD",
  "Status 3: IABP",
  "Status 4: LVAD"
)

#generate numerators for each criteria
criteria_month_props <- criteria_numerators %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = all_patient_monthly_counts$MonthSnapshot,
    support_type = support_levels,
    fill = list(n_yes = 0)
  ) %>%
  left_join(all_patient_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_all
  )
```

## Overall support-type proportions

```{r overall-support-proportions}
## Overall support-type proportions

overall_support_month_props <- criteria_numerators %>%
  mutate(
    support_type = case_when(
      support_type == "Status 1: ECMO" ~ "ECMO",
      support_type %in% c("Status 2: IABP", "Status 3: IABP") ~ "IABP",
      support_type == "Status 2: Endovascular MCSD" ~ "Endovascular MCSD",
      support_type == "Status 4: LVAD" ~ "LVAD",
      TRUE ~ support_type
    )
  ) %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = all_patient_monthly_counts$MonthSnapshot,
    support_type = c("ECMO", "IABP", "Endovascular MCSD", "LVAD"),
    fill = list(n_yes = 0)
  ) %>%
  left_join(all_patient_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_all
  )
```

```{r overall-support-plot}
ggplot(
  overall_support_month_props,
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.50),
    breaks = seq(0, 0.50, 0.05),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Active Heart Candidates by Support Type",
    x = "Month Snapshot",
    y = "Proportion of Active Status 1–6 Candidates",
    color = "Support Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Same graphs w/o lvad on there

```{r}
ggplot(
  overall_support_month_props %>%
    filter(support_type != "LVAD"),
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.10),
    breaks = seq(0, 0.10, 0.01),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Active Heart Candidates by Support Type",
    subtitle = "LVAD excluded",
    x = "Month Snapshot",
    y = "Proportion of Active Status 1–6 Candidates",
    color = "Support Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


## Status 2 devices and exceptions

These are Status 2-related numerators, but the denominator is still `all_patient_monthly_counts` (NOT JUST STATUS 2)

```{r status2-device-exception-proportions}
# Active Status 2 patient-month denominator from stathist
active_status2_patient_months <- stathist_monthly %>%
  filter(CANHX_STAT_label == "Adult Status 2") %>%
  filter(!is.na(PX_ID)) %>%
  distinct(PX_ID, MonthSnapshot)

status2_monthly_counts <- active_status2_patient_months %>%
  group_by(MonthSnapshot) %>%
  summarise(
    n_status2 = n_distinct(PX_ID),
    .groups = "drop"
  )

# Keep only deduped JustForm patient-months that are actually active Status 2
# in stathist on the same MonthSnapshot
status2_patient_month_criteria <- justform_patient_month %>%
  filter(status_form == "Adult Status 2")

# Status 2 device and exception numerators
status2_device_exception_numerators <- bind_rows(
  status2_patient_month_criteria %>%
    group_by(MonthSnapshot) %>%
    summarise(
      support_type = "Status 2: IABP",
      n_yes = sum(iabp, na.rm = TRUE),
      .groups = "drop"
    ),
  
  status2_patient_month_criteria %>%
    group_by(MonthSnapshot) %>%
    summarise(
      support_type = "Status 2: Endovascular MCSD",
      n_yes = sum(endovascular_mcsd, na.rm = TRUE),
      .groups = "drop"
    ),
  
  status2_patient_month_criteria %>%
    group_by(MonthSnapshot) %>%
    summarise(
      support_type = "Status 2: Exception",
      n_yes = sum(exception, na.rm = TRUE),
      .groups = "drop"
    )
)

# Final Status 2 proportions
status2_device_exception_props <- status2_device_exception_numerators %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = status2_monthly_counts$MonthSnapshot,
    support_type = c(
      "Status 2: IABP",
      "Status 2: Endovascular MCSD",
      "Status 2: Exception"
    ),
    fill = list(n_yes = 0)
  ) %>%
  left_join(status2_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_status2
  )
```

```{r status2-device-exception-plot}
ggplot(
  status2_device_exception_props,
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(
    limits = c(0, 0.8),
    breaks = seq(0, 0.8, 0.08),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Status 2 Candidates by Device/Exception Criterion",
    x = "Month Snapshot",
    y = "Proportion of Active Status 2 Candidates",
    color = "Criterion"
  ) +
  theme_minimal()
```

```{r status2-device-exception-plot-2025}
ggplot(
  status2_device_exception_props %>%
    filter(MonthSnapshot >= as.Date("2025-01-01")),
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.8),
    breaks = seq(0, 0.8, 0.08),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Status 2 Candidates by Device/Exception Criterion, 2025 Onward",
    x = "Month Snapshot",
    y = "Proportion of Active Status 2 Candidates",
    color = "Criterion"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r}
#checking that no one has both iabp AND endo_mcsd
status2_patient_month_criteria %>%
  summarise(
    n_iabp = sum(iabp),
    n_endo_mcsd = sum(endovascular_mcsd),
    n_both = sum(iabp & endovascular_mcsd),
    n_either = sum(iabp | endovascular_mcsd),
    n_iabp_plus_endo = sum(iabp) + sum(endovascular_mcsd)
  )

#making new snapshot var for iabp OR endo_mcsd
status2_iabp_endo_monthly <- status2_patient_month_criteria %>%
  group_by(MonthSnapshot) %>%
  summarise(
    support_type = "Status 2: IABP or Endovascular MCSD",
    n_yes = n_distinct(PX_ID[iabp | endovascular_mcsd]),
    .groups = "drop"
  ) %>%
  left_join(status2_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_status2
  )

#pulling exceptions variable
status2_exception_monthly <- status2_device_exception_props %>%
  filter(support_type == "Status 2: Exception") %>%
  select(MonthSnapshot, support_type, n_yes, n_status2, proportion)

#combined dataset for plotting
status2_iabp_endo_vs_exception <- bind_rows(
  status2_iabp_endo_monthly,
  status2_exception_monthly
)

#plots
ggplot(
  status2_iabp_endo_vs_exception,
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(
    limits = c(0, 0.8),
    breaks = seq(0, 0.8, 0.08),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Status 2 Candidates: IABP/Endovascular MCSD vs Exceptions",
    x = "Month Snapshot",
    y = "Proportion of Active Status 2 Candidates",
    color = "Criterion"
  ) +
  theme_minimal()

ggplot(
  status2_iabp_endo_vs_exception %>%
    filter(MonthSnapshot >= as.Date("2025-01-01")),
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.8),
    breaks = seq(0, 0.8, 0.08),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Status 2 Candidates: IABP/Endovascular MCSD vs Exceptions, 2025 Onward",
    x = "Month Snapshot",
    y = "Proportion of Active Status 2 Candidates",
    color = "Criterion"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )



```

