---
title: "JustForms"
output:
  html_document:
    self_contained: true
date: "2026-06-16"
---

```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)

# Snapshot window shortened to 2024-09-01 .. 2026-03-01.
snapshot_start <- as.Date("2024-09-01")
snapshot_end   <- as.Date("2026-03-01")
policy_date    <- as.Date("2025-09-01")
```

## Load data

```{r load-data}
JustFormHRStat1 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat1.sas7bdat")
JustFormHRStat2 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat2.sas7bdat")
JustFormHRStat3 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat3.sas7bdat")
JustFormHRStat4 <- read_sas("/Users/chrisramos/Downloads/JustFormHRStat4.sas7bdat")
JustFormHR      <- read_sas("/Users/chrisramos/Downloads/JustFormHR.sas7bdat")
StatHist_Thor   <- read_sas("/Users/chrisramos/Downloads/stathist_thor.sas7bdat")
JustFormHRDataLink <- read_sas("/Users/chrisramos/Downloads/JustFormHRDataLink.sas7bdat")
RiskStratDataHR <- read_sas("/Users/chrisramos/Downloads/RiskStratDataHR.sas7bdat")

View(stathist_thor)
View(JustFormHRStat1)
```

##Linking Just_ID to PX_ID

```{r}
justid_px_lookup <- JustFormHRDataLink %>%    #this dataset links JustID and WlregAuditID
  mutate(
    JustId = as.character(JustId),
    WlregAuditId = as.character(WlregAuditId)
  ) %>%

  select(JustId, WlregAuditId) %>%
  left_join(
    RiskStratDataHR %>%                      #this dataset link WlregAuditID to PX_ID
      mutate(
        WlregAuditId = as.character(WlregAuditId)
      ) %>%
      select(WlregAuditId, px_id),
    by = "WlregAuditId"                      #join our two simplified tables by common waitlist reg ID
  ) %>%
  rename(PX_ID = px_id)

#justid_px_lookup now has all 3 IDs joined by common Wlreg ID; If several JustIds share one WlregAuditId, they all get the same PX_ID

#sanity checks

#shows each just_id has a specific wlregaudit_id accounted for by a each row; as expected some patients have mult. of these ids
justid_px_lookup %>%
  summarise(
    n_rows = n(),
    n_JustId = n_distinct(JustId),
    n_WlregAuditId = n_distinct(WlregAuditId),
    n_PX_ID = n_distinct(PX_ID),
    missing_px_id = sum(is.na(PX_ID))
  )

#there are no duplicated just_ids
justid_px_lookup %>%
  count(JustId) %>%
  filter(n > 1)

#checking how many just_ids each px_id has associated with them
justid_px_lookup %>%
  group_by(PX_ID) %>%
  summarise(
    n_JustId = n_distinct(JustId),
    n_WlregAuditId = n_distinct(WlregAuditId),
    .groups = "drop"
  ) %>%
  arrange(desc(n_JustId))
```


## Create monthly snapshots

```{r monthly-snapshots}
## Monthly snapshot dates now come from status history, not JustForms

month_checkpoints <- seq.Date(
  from = snapshot_start,
  to = snapshot_end,
  by = "month"
)

add_active_month_snapshots <- function(df) {
  df %>%
    mutate(
      CANHX_BEGIN_DT = as.Date(CANHX_BEGIN_DT),
      CANHX_END_DT = as.Date(CANHX_END_DT),
      CAN_REM_DT = as.Date(CAN_REM_DT),
      CANHX_STAT_CD = as.character(CANHX_STAT_CD),
      
      # If no status-history end date exists, use removal date.
      # If neither exists, assume the status persists through snapshot_end.
      CANHX_END_DT_adjusted = case_when(
        !is.na(CANHX_END_DT) ~ CANHX_END_DT,
        is.na(CANHX_END_DT) & !is.na(CAN_REM_DT) ~ CAN_REM_DT,
        is.na(CANHX_END_DT) & is.na(CAN_REM_DT) ~ snapshot_end
      )
    ) %>%
    filter(
      CANHX_BEGIN_DT <= snapshot_end,
      CANHX_END_DT_adjusted >= snapshot_start
    ) %>%
    rowwise() %>%
    mutate(
      MonthSnapshot = list(
        month_checkpoints[
          month_checkpoints >= CANHX_BEGIN_DT &
            month_checkpoints <= CANHX_END_DT_adjusted
        ]
      )
    ) %>%
    ungroup() %>%
    tidyr::unnest(MonthSnapshot)
}

## No longer creating MonthSnapshot for JustForms. Using ChangeDate so later we can ask for most recent JustForm as of MonthSnapshot.
JustFormHRStat1 <- JustFormHRStat1 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat2 <- JustFormHRStat2 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat3 <- JustFormHRStat3 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHRStat4 <- JustFormHRStat4 %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

JustFormHR <- JustFormHR %>%
  mutate(
    JustId = as.character(JustId),
    ChangeDate = as.Date(ChangeDate)
  )

#clean JustId -> PX_ID lookup
justid_px_lookup_clean <- justid_px_lookup %>%
  mutate(
    JustId = as.character(JustId),
    PX_ID = as.character(PX_ID)
  ) %>%
  select(JustId, PX_ID) %>%
  distinct(JustId, .keep_all = TRUE)

#clean JustForm core
justform_core <- JustFormHR %>%
  transmute(
    JustId,
    RequestedCandStat_descrip,
    exception = Exception == 1
  ) %>%
  distinct(JustId, .keep_all = TRUE)

#build long JustForm table; keeps all forms up to snapshot_end. The most-recent-form-as-of-MonthSnapshot logic (below) will pick the governing form per month, so we retain forms dated before snapshot_start too (a form from, say, early 2024 can still be the most recent one governing 2024-09).
justform_status_long <- bind_rows(
  
  JustFormHRStat1 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 1",
      ecmo = CriteriaEcmoSupport == 1,
      iabp = FALSE,
      endovascular_mcsd = FALSE,
      lvad = FALSE,
      lvad_disc = FALSE,        # Status 3 only; FALSE here so bind_rows aligns
      inotrope_high = FALSE,    # Status 3 only
      inotrope_low = FALSE      # Status 4 only
    ),
  
  JustFormHRStat2 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 2",
      ecmo = FALSE,
      iabp = CriteriaIabpSupport == 1,
      endovascular_mcsd = CriteriaMcsdEndovasSupp == 1,   # impella == endovascular MCSD (Status 2 only)
      lvad = FALSE,
      lvad_disc = FALSE,
      inotrope_high = FALSE,
      inotrope_low = FALSE
    ),
  
  JustFormHRStat3 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 3",
      ecmo = FALSE,
      iabp = CriteriaIabpSupport == 1,
      endovascular_mcsd = FALSE,
      lvad = FALSE,
      lvad_disc = CriteriaLvadDiscSupport == 1,            # LVAD discretionary (Status 3)
      inotrope_high = CriteriaInotropeSupport == 1,        # high-dose inotropes (Status 3 file)
      inotrope_low = FALSE
    ),
  
  JustFormHRStat4 %>%
    transmute(
      JustId,
      ChangeDate,
      status_form = "Adult Status 4",
      ecmo = FALSE,
      iabp = FALSE,
      endovascular_mcsd = FALSE,
      lvad = CriteriaLvadSupport == 1,
      lvad_disc = FALSE,
      inotrope_high = FALSE,
      inotrope_low = CriteriaInotropeSupport == 1          # low-dose inotropes (Status 4 file)
    )
) %>%
  filter(
    !is.na(ChangeDate),
    ChangeDate <= snapshot_end
  ) %>%
  left_join(
    justid_px_lookup_clean,
    by = "JustId"
  ) %>%
  left_join(
    justform_core,
    by = "JustId"
  ) %>%
  mutate(
    PX_ID = as.character(PX_ID)
  )

```

##Prepare stathist_thor and get a total monthly candidate count across all 6 statuses
```{r}
## stathist_thor monthly active denominator

#heart-only status history rows, expanded to active first-of-month snapshots
stathist_filtered <- StatHist_Thor %>%
  mutate(
    PX_ID = as.character(PX_ID)
  ) %>%
  filter(WL_ORG == "HR") %>%
  add_active_month_snapshots()

#keep adult statuses 1-6 and label them
stathist_monthly <- stathist_filtered %>%
  filter(CANHX_STAT_CD %in% c(
    "2110", "2120", "2130", "2140", "2150", "2160"
  )) %>%
  mutate(
    CANHX_STAT_label = recode(
      CANHX_STAT_CD,
      "2110" = "Adult Status 1",
      "2120" = "Adult Status 2",
      "2130" = "Adult Status 3",
      "2140" = "Adult Status 4",
      "2150" = "Adult Status 5",
      "2160" = "Adult Status 6"
    )
  )

#gets patient/status/month frame from stathist
active_status_patient_months <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  distinct(
    PX_ID,
    MonthSnapshot,
    status_form = CANHX_STAT_label
  )

#for each active patient/status/month, take the most recent JustForm
justform_patient_month <- active_status_patient_months %>%
  inner_join(
    justform_status_long %>%
      filter(
        !is.na(PX_ID),
        !is.na(ChangeDate)
      ) %>%
      select(
        PX_ID,
        JustId,
        ChangeDate,
        status_form,
        RequestedCandStat_descrip,
        ecmo,
        iabp,
        endovascular_mcsd,
        lvad,
        lvad_disc,
        inotrope_high,
        inotrope_low,
        exception
      ),
    by = c("PX_ID", "status_form"),
    relationship = "many-to-many"
  ) %>%
  filter(
    ChangeDate <= MonthSnapshot
  ) %>%
  group_by(PX_ID, MonthSnapshot, status_form) %>%
  arrange(
    desc(ChangeDate),
    desc(suppressWarnings(as.numeric(JustId))),
    .by_group = TRUE
  ) %>%
  slice(1) %>%
  ungroup()

#check that there is only one JustForm row per patient/status/month.
justform_patient_month %>%
  count(PX_ID, MonthSnapshot, status_form) %>%
  filter(n > 1)
```


## Total monthly denominator (total candidates listed per month (stat 1-6))

```{r total-denominator}
#Total monthly denominator: unique active Status 1-6 heart candidates
all_patient_monthly_counts <- stathist_monthly %>%
  filter(!is.na(PX_ID)) %>%
  group_by(MonthSnapshot) %>%
  summarise(
    n_all = n_distinct(PX_ID),
    .groups = "drop"
  )
```

## Monthly numerators and proportions (how much of each device)

```{r monthly-numerators}
#how many (individuals, month) have criterion of interest (adds new label, keeps status value as well)
summarise_patient_criterion <- function(status_value, flag_col, support_label) {
  justform_patient_month %>%
    group_by(MonthSnapshot) %>%
    summarise(
      support_type = support_label,
      n_yes = sum(
        status_form == status_value & .data[[flag_col]],
        na.rm = TRUE
      ),
      .groups = "drop"
    )
}

#use function to generate all numerators desired + label
criteria_numerators <- bind_rows(
  summarise_patient_criterion("Adult Status 1", "ecmo", "Status 1: ECMO"),
  summarise_patient_criterion("Adult Status 2", "iabp", "Status 2: IABP"),
  summarise_patient_criterion("Adult Status 2", "endovascular_mcsd", "Status 2: Endovascular MCSD"),
  summarise_patient_criterion("Adult Status 3", "iabp", "Status 3: IABP"),
  summarise_patient_criterion("Adult Status 4", "lvad", "Status 4: LVAD")
)

support_levels <- c(
  "Status 1: ECMO",
  "Status 2: IABP",
  "Status 2: Endovascular MCSD",
  "Status 3: IABP",
  "Status 4: LVAD"
)

#generate numerators for each criteria
criteria_month_props <- criteria_numerators %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = all_patient_monthly_counts$MonthSnapshot,
    support_type = support_levels,
    fill = list(n_yes = 0)
  ) %>%
  left_join(all_patient_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_all
  )
```

## Overall support-type proportions

```{r overall-support-proportions}
overall_support_month_props <- criteria_numerators %>%
  mutate(
    support_type = case_when(
      support_type == "Status 1: ECMO" ~ "ECMO",
      support_type %in% c("Status 2: IABP", "Status 3: IABP") ~ "IABP",
      support_type == "Status 2: Endovascular MCSD" ~ "Endovascular MCSD",
      support_type == "Status 4: LVAD" ~ "LVAD",
      TRUE ~ support_type
    )
  ) %>%
  group_by(MonthSnapshot, support_type) %>%
  summarise(
    n_yes = sum(n_yes),
    .groups = "drop"
  ) %>%
  complete(
    MonthSnapshot = all_patient_monthly_counts$MonthSnapshot,
    support_type = c("ECMO", "IABP", "Endovascular MCSD", "LVAD"),
    fill = list(n_yes = 0)
  ) %>%
  left_join(all_patient_monthly_counts, by = "MonthSnapshot") %>%
  mutate(
    proportion = n_yes / n_all
  )
```

```{r overall-support-plot}
ggplot(
  overall_support_month_props,
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.50),
    breaks = seq(0, 0.50, 0.05),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Active Heart Candidates by Support Type",
    x = "Month Snapshot",
    y = "Proportion of Active Status 1–6 Candidates",
    color = "Support Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Same graphs w/o lvad on there

```{r}
ggplot(
  overall_support_month_props %>%
    filter(support_type != "LVAD"),
  aes(x = MonthSnapshot, y = proportion, color = support_type)
) +
  geom_line() +
  geom_point(size = 1) +
  geom_vline(xintercept = policy_date, linetype = "dashed") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
  scale_y_continuous(
    limits = c(0, 0.10),
    breaks = seq(0, 0.10, 0.01),
    labels = scales::percent_format()
  ) +
  labs(
    title = "Monthly Proportion of Active Heart Candidates by Support Type",
    subtitle = "LVAD excluded",
    x = "Month Snapshot",
    y = "Proportion of Active Status 1–6 Candidates",
    color = "Support Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```


## Within-status numerator helper

```{r within-status-helper}
# Reusable within-status builders. Pattern used by ALL of Status 1-4 below:
#   numerator   = active-in-status patient-months whose governing JustForm has the criterion of interest
#   denominator = active-in-status patient-months from stathist
# Every status section uses the same three helpers and the same plot styling, so the
# graphs are directly comparable.
#
# Returns (for each helper): MonthSnapshot, support_type, n_yes, n_status, proportion.

# denominator for any status, from stathist
status_monthly_denominator <- function(status_label) {
  stathist_monthly %>%
    filter(CANHX_STAT_label == status_label, !is.na(PX_ID)) %>%
    distinct(PX_ID, MonthSnapshot) %>%
    group_by(MonthSnapshot) %>%
    summarise(n_status = n_distinct(PX_ID), .groups = "drop")
}

# one within-status criterion: distinct patients whose governing form has `flag_col`
within_status_criterion <- function(criteria_rows, denom, flag_col, support_label) {
  criteria_rows %>%
    group_by(MonthSnapshot) %>%
    summarise(
      n_yes = n_distinct(PX_ID[.data[[flag_col]]]),
      .groups = "drop"
    ) %>%
    complete(MonthSnapshot = denom$MonthSnapshot, fill = list(n_yes = 0)) %>%
    mutate(support_type = support_label) %>%
    left_join(denom, by = "MonthSnapshot") %>%
    mutate(proportion = n_yes / n_status) %>%
    select(MonthSnapshot, support_type, n_yes, n_status, proportion)
}

# within-status count over an arbitrary row-predicate (used for "Other" and for combined
# device = IABP or impella). `predicate_fn` is a function of the data frame returning a
# logical vector (e.g. function(d) d$iabp | d$endovascular_mcsd); distinct patients
# satisfying it are counted. Using a plain function avoids any tidy-eval/rlang fragility.
within_status_union <- function(criteria_rows, denom, predicate_fn, support_label) {
  df <- criteria_rows
  df$.match <- predicate_fn(df)        # logical vector over all rows, computed up front
  df %>%
    group_by(MonthSnapshot) %>%
    summarise(
      n_yes = n_distinct(PX_ID[.match]),
      .groups = "drop"
    ) %>%
    complete(MonthSnapshot = denom$MonthSnapshot, fill = list(n_yes = 0)) %>%
    mutate(support_type = support_label) %>%
    left_join(denom, by = "MonthSnapshot") %>%
    mutate(proportion = n_yes / n_status) %>%
    select(MonthSnapshot, support_type, n_yes, n_status, proportion)
}

# shared plot styling so every within-status graph matches
within_status_plot <- function(df, plot_title, y_label = "Proportion of Active Candidates",
                                y_max = 0.8, y_step = 0.08) {
  ggplot(df, aes(x = MonthSnapshot, y = proportion, color = support_type)) +
    geom_line() +
    geom_point(size = 1) +
    geom_vline(xintercept = policy_date, linetype = "dashed") +
    scale_x_date(date_breaks = "1 month", date_labels = "%m/%Y") +
    scale_y_continuous(
      limits = c(0, y_max),
      breaks = seq(0, y_max, y_step),
      labels = scales::percent_format()
    ) +
    labs(title = plot_title, x = "Month Snapshot", y = y_label, color = "Criterion") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


## Status 4: LVAD vs low-dose inotropes (within Status 4)

```{r status4-criteria-proportions}
status4_denom <- status_monthly_denominator("Adult Status 4")

status4_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 4")

status4_props <- bind_rows(
  within_status_criterion(status4_criteria_rows, status4_denom, "lvad",         "Status 4: LVAD"),
  within_status_criterion(status4_criteria_rows, status4_denom, "inotrope_low", "Status 4: Low-dose inotropes")
)

status4_props
```

```{r status4-criteria-plot}
within_status_plot(
  status4_props,
  "Monthly Proportion of Status 4 Candidates by Criterion",
  y_label = "Proportion of Active Status 4 Candidates"
)
```


## Status 3: LVAD discretionary vs high-dose inotropes (within Status 3)

LVAD complications criterion is deferred for now (to be added later).

```{r status3-criteria-proportions}
status3_denom <- status_monthly_denominator("Adult Status 3")

status3_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 3")

status3_props <- bind_rows(
  within_status_criterion(status3_criteria_rows, status3_denom, "lvad_disc",     "Status 3: LVAD discretionary"),
  within_status_criterion(status3_criteria_rows, status3_denom, "inotrope_high", "Status 3: High-dose inotropes")
  # TODO: add "Status 3: LVAD complications" once that criterion source is identified
)

status3_props
```

```{r status3-criteria-plot}
within_status_plot(
  status3_props,
  "Monthly Proportion of Status 3 Candidates by Criterion",
  y_label = "Proportion of Active Status 3 Candidates"
)
```


## Status 2: IABP, impella, exceptions, and Other (within Status 2)

```{r status2-overlap-check}
status2_denom <- status_monthly_denominator("Adult Status 2")

status2_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 2")

status2_overlap_check <- status2_criteria_rows %>%
  summarise(
    n_rows                 = n(),
    n_iabp                 = sum(iabp, na.rm = TRUE),
    n_impella_endovas      = sum(endovascular_mcsd, na.rm = TRUE),
    n_exception            = sum(exception, na.rm = TRUE),
    n_device               = sum((iabp | endovascular_mcsd), na.rm = TRUE),
    n_device_and_exception = sum((iabp | endovascular_mcsd) & exception, na.rm = TRUE),
    n_iabp_and_impella     = sum(iabp & endovascular_mcsd, na.rm = TRUE)
  )

status2_overlap_check
```

```{r status2-broken-out-proportions}
# (A) IABP / impella / exception / other, broken out
status2_props_brokenout <- bind_rows(
  within_status_criterion(status2_criteria_rows, status2_denom, "iabp",              "Status 2: IABP"),
  within_status_criterion(status2_criteria_rows, status2_denom, "endovascular_mcsd", "Status 2: Impella (Endovascular MCSD)"),
  within_status_criterion(status2_criteria_rows, status2_denom, "exception",         "Status 2: Exception"),
  within_status_union(status2_criteria_rows, status2_denom,
                      function(d) !d$exception & !d$iabp & !d$endovascular_mcsd, "Status 2: Other")
)

status2_props_brokenout
```

```{r status2-broken-out-plot}
within_status_plot(
  status2_props_brokenout,
  "Monthly Proportion of Status 2 Candidates by Criterion (broken out)",
  y_label = "Proportion of Active Status 2 Candidates"
)
```

```{r status2-combined-proportions}
# (B) combined device (IABP or impella) / exception / other
status2_props_combined <- bind_rows(
  within_status_union(status2_criteria_rows, status2_denom,
                      function(d) d$iabp | d$endovascular_mcsd, "Status 2: IABP or Impella"),
  within_status_criterion(status2_criteria_rows, status2_denom, "exception", "Status 2: Exception"),
  within_status_union(status2_criteria_rows, status2_denom,
                      function(d) !d$exception & !d$iabp & !d$endovascular_mcsd, "Status 2: Other")
)

status2_props_combined
```

```{r status2-combined-plot}
within_status_plot(
  status2_props_combined,
  "Monthly Proportion of Status 2 Candidates: Device vs Exception vs Other",
  y_label = "Proportion of Active Status 2 Candidates"
)
```


## Status 1: ECMO, exceptions, and Other (within Status 1)

"Other" = active Status 1 patients whose governing form is neither ECMO nor exception.

```{r status1-overlap-check}
status1_denom <- status_monthly_denominator("Adult Status 1")

status1_criteria_rows <- justform_patient_month %>%
  filter(status_form == "Adult Status 1")

status1_overlap_check <- status1_criteria_rows %>%
  summarise(
    n_rows               = n(),
    n_ecmo               = sum(ecmo, na.rm = TRUE),
    n_exception          = sum(exception, na.rm = TRUE),
    n_ecmo_and_exception = sum(ecmo & exception, na.rm = TRUE),
    n_ecmo_or_exception  = sum(ecmo | exception, na.rm = TRUE)
  )

status1_overlap_check
```

```{r status1-proportions}
status1_props <- bind_rows(
  within_status_criterion(status1_criteria_rows, status1_denom, "ecmo",      "Status 1: ECMO"),
  within_status_criterion(status1_criteria_rows, status1_denom, "exception", "Status 1: Exception"),
  within_status_union(status1_criteria_rows, status1_denom,
                      function(d) !d$ecmo & !d$exception, "Status 1: Other")
)

status1_props
```

```{r status1-plot}
within_status_plot(
  status1_props,
  "Monthly Proportion of Status 1 Candidates by Criterion",
  y_label = "Proportion of Active Status 1 Candidates"
)
```

```{r}

# justform_patient_month is an INNER join of stathist active patient-months with the
# justification forms, so it only contains patient-months that have a governing form
# (ChangeDate <= MonthSnapshot) whose form-status matches the stathist status. The
# within-status denominators count ALL active patients in that status from stathist,
# including those with no governing form that month -- which is why IABP + impella +
# exception + Other falls short of 1. This chunk quantifies and characterizes the gap.

# most-recent form per (PX_ID, MonthSnapshot) across ANY status (not status-matched),
# so we can tell "no same-status form" apart from "no form at all".
most_recent_any_status_form <- active_status_patient_months %>%
  distinct(PX_ID, MonthSnapshot) %>%
  inner_join(
    justform_status_long %>%
      filter(!is.na(PX_ID), !is.na(ChangeDate)) %>%
      select(PX_ID, JustId, ChangeDate, status_form),
    by = "PX_ID",
    relationship = "many-to-many"
  ) %>%
  filter(ChangeDate <= MonthSnapshot) %>%
  group_by(PX_ID, MonthSnapshot) %>%
  arrange(desc(ChangeDate), desc(suppressWarnings(as.numeric(JustId))), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  transmute(PX_ID, MonthSnapshot, any_form_status = status_form)

# per-status form coverage summary (pooled over all months in window)
form_coverage_by_status <- lapply(
  c("Adult Status 1", "Adult Status 2", "Adult Status 3",
    "Adult Status 4", "Adult Status 5", "Adult Status 6"),
  function(stat) {
    active <- active_status_patient_months %>%
      filter(status_form == stat) %>%
      distinct(PX_ID, MonthSnapshot)

    has_matched_form <- justform_patient_month %>%
      filter(status_form == stat) %>%
      distinct(PX_ID, MonthSnapshot) %>%
      mutate(has_matched_form = TRUE)

    active %>%
      left_join(has_matched_form, by = c("PX_ID", "MonthSnapshot")) %>%
      mutate(has_matched_form = !is.na(has_matched_form)) %>%
      left_join(most_recent_any_status_form, by = c("PX_ID", "MonthSnapshot")) %>%
      summarise(
        status_form          = stat,
        n_active             = n(),
        n_matched_form       = sum(has_matched_form),
        n_no_matched_form    = sum(!has_matched_form),
        n_form_other_status  = sum(!has_matched_form & !is.na(any_form_status)),
        n_no_form_at_all     = sum(!has_matched_form & is.na(any_form_status)),
        .groups = "drop"
      ) %>%
      mutate(
        pct_matched = round(100 * n_matched_form / n_active, 1),
        pct_no_form = round(100 * n_no_matched_form / n_active, 1)
      )
  }
) %>% bind_rows()

form_coverage_by_status

# Status 2 broken out by month so you can see if the gap is steady or concentrated
# (e.g. at the left edge of the window).
cat("\n=== STATUS 2 FORM COVERAGE BY MONTH ===\n")
status2_coverage_by_month <- active_status_patient_months %>%
  filter(status_form == "Adult Status 2") %>%
  distinct(PX_ID, MonthSnapshot) %>%
  left_join(
    justform_patient_month %>%
      filter(status_form == "Adult Status 2") %>%
      distinct(PX_ID, MonthSnapshot) %>%
      mutate(has_matched_form = TRUE),
    by = c("PX_ID", "MonthSnapshot")
  ) %>%
  mutate(has_matched_form = !is.na(has_matched_form)) %>%
  group_by(MonthSnapshot) %>%
  summarise(
    n_active       = n(),
    n_matched_form = sum(has_matched_form),
    n_no_form      = sum(!has_matched_form),
    pct_no_form    = round(100 * sum(!has_matched_form) / n(), 1),
    .groups = "drop"
  )

status2_coverage_by_month
```

```{r}
# All active patient-months (stathist truth)
active_pm <- active_status_patient_months %>%
  distinct(PX_ID, MonthSnapshot, status_form)

# Matched ones (survived the inner join + ChangeDate<=MonthSnapshot)
matched_pm <- justform_patient_month %>%
  distinct(PX_ID, MonthSnapshot, status_form) %>%
  mutate(matched = TRUE)

# Every form for a patient, with its filed status and date (no status/date filter yet)
all_forms <- justform_status_long %>%
  filter(!is.na(PX_ID), !is.na(ChangeDate)) %>%
  select(PX_ID, JustId, ChangeDate, form_status = status_form)

unmatched <- active_pm %>%
  left_join(matched_pm, by = c("PX_ID", "MonthSnapshot", "status_form")) %>%
  filter(is.na(matched)) %>%
  select(PX_ID, MonthSnapshot, listed_status = status_form)

# For each unmatched (PX_ID, month, listed_status), characterize WHY:
mismatch_anatomy <- unmatched %>%
  left_join(all_forms, by = "PX_ID", relationship = "many-to-many") %>%
  group_by(PX_ID, MonthSnapshot, listed_status) %>%
  summarise(
    has_any_form_ever              = any(!is.na(JustId)),
    has_any_form_by_month          = any(ChangeDate <= MonthSnapshot),
    has_same_status_form_ever      = any(form_status == listed_status),
    has_same_status_form_by_month  = any(form_status == listed_status & ChangeDate <= MonthSnapshot),
    # most recent form AS OF the month, regardless of status
    recent_form_status_by_month    = form_status[ChangeDate <= MonthSnapshot][which.max(ChangeDate[ChangeDate <= MonthSnapshot])][1],
    .groups = "drop"
  ) %>%
  mutate(
    reason = case_when(
      !has_any_form_ever                              ~ "1. no form in any StatN file at all",
      !has_any_form_by_month                          ~ "3b. only has form(s) dated AFTER this month",
      has_same_status_form_by_month                   ~ "?? same-status form exists by month (should have matched - investigate)",
      !has_same_status_form_ever                      ~ "2. has form(s) but NEVER in the listed status",
      TRUE                                            ~ "3a. has same-status form, but only dated after month; most recent by-month is another status"
    )
  )

cat("\n=== WHY UNMATCHED (counts by reason) ===\n")
mismatch_anatomy %>% count(reason, sort = TRUE) %>% print()

cat("\n=== for status mismatches: listed status vs the most-recent by-month form status ===\n")
mismatch_anatomy %>%
  filter(grepl("^2|^3a", reason)) %>%
  count(listed_status, recent_form_status_by_month, sort = TRUE) %>%
  print(n = 40)

cat("\n=== sample of the worst category for manual inspection ===\n")
mismatch_anatomy %>% filter(grepl("investigate", reason)) %>% head(20) %>% print()
```

```{r}
# 1. How many raw justification forms FAIL to resolve to a PX_ID?
#    (justform_status_long before the !is.na(PX_ID) filter.)
raw_forms_linked <- bind_rows(
  JustFormHRStat1 %>% transmute(JustId, status_form = "Adult Status 1"),
  JustFormHRStat2 %>% transmute(JustId, status_form = "Adult Status 2"),
  JustFormHRStat3 %>% transmute(JustId, status_form = "Adult Status 3"),
  JustFormHRStat4 %>% transmute(JustId, status_form = "Adult Status 4")
) %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(justid_px_lookup_clean, by = "JustId")

cat("\n=== RAW STATN FORM LINKAGE ===\n")
raw_forms_linked %>%
  summarise(
    n_forms              = n(),
    n_forms_with_PX_ID   = sum(!is.na(PX_ID)),
    n_forms_missing_PX_ID = sum(is.na(PX_ID)),
    pct_missing          = round(100 * mean(is.na(PX_ID)), 2)
  ) %>% print()

# 2. Are the JustIds that fail to link actually present in the DataLink table at all?
#    distinguishes "JustId absent from DataLink" vs "DataLink row maps to NA px_id"
cat("\n=== WHERE DOES THE LINK BREAK? ===\n")
unlinked_justids <- raw_forms_linked %>% filter(is.na(PX_ID)) %>% distinct(JustId)
unlinked_justids %>%
  left_join(
    JustFormHRDataLink %>% mutate(JustId = as.character(JustId)) %>%
      distinct(JustId, WlregAuditId = as.character(WlregAuditId)),
    by = "JustId"
  ) %>%
  summarise(
    n_unlinked_justids        = n(),
    n_absent_from_datalink    = sum(is.na(WlregAuditId)),       # JustId not in DataLink at all
    n_in_datalink_but_no_pxid = sum(!is.na(WlregAuditId))       # in DataLink, but WlregAuditId->px_id failed
  ) %>% print()

# 3. Reconcile to bucket 1: do bucket-1 PX_IDs gain a form if we DON'T require linkage?
#    We can only attach unlinked forms to a PX_ID through WlregAuditId->stathist if that
#    path exists; if not, these forms are simply unattributable.
cat("\nNote: if n_forms_missing_PX_ID is ~0, bucket 1 is genuinely form-less (no fix).\n")
cat("If it is large, we are dropping real forms at the JustId->PX_ID bridge (fixable).\n")
```
```{r}
# Bucket 1 ("no form in any StatN file") only means a PX_ID has no form that LINKED to it.
# That could be (a) the patient genuinely never had a justification form, or (b) a form
# exists in the raw StatN file but its JustId failed to resolve to a PX_ID through the chain
# JustId -> JustFormHRDataLink -> WlregAuditId -> RiskStratDataHR -> px_id, so we dropped it
# with !is.na(PX_ID). This chunk measures how many raw forms fail to link, and WHERE.

# 1. Of all RAW StatN forms, how many resolve to a PX_ID vs not?
raw_forms_linked <- bind_rows(
  JustFormHRStat1 %>% transmute(JustId, status_form = "Adult Status 1"),
  JustFormHRStat2 %>% transmute(JustId, status_form = "Adult Status 2"),
  JustFormHRStat3 %>% transmute(JustId, status_form = "Adult Status 3"),
  JustFormHRStat4 %>% transmute(JustId, status_form = "Adult Status 4")
) %>%
  mutate(JustId = as.character(JustId)) %>%
  left_join(justid_px_lookup_clean, by = "JustId")

cat("\n=== RAW STATN FORM LINKAGE ===\n")
raw_forms_linked %>%
  summarise(
    n_forms               = n(),
    n_forms_with_PX_ID    = sum(!is.na(PX_ID)),
    n_forms_missing_PX_ID = sum(is.na(PX_ID)),
    pct_missing           = round(100 * mean(is.na(PX_ID)), 2)
  ) %>% print()

# 2. For forms that fail to link: JustId absent from DataLink, or present but
#    WlregAuditId -> px_id yields NA, or actually resolvable (=> our lookup has a bug).
cat("\n=== WHERE DOES THE LINK BREAK? ===\n")
datalink_min <- JustFormHRDataLink %>%
  mutate(JustId = as.character(JustId), WlregAuditId = as.character(WlregAuditId)) %>%
  distinct(JustId, WlregAuditId)

riskstrat_min <- RiskStratDataHR %>%
  mutate(WlregAuditId = as.character(WlregAuditId)) %>%
  distinct(WlregAuditId, px_id)

raw_forms_linked %>%
  filter(is.na(PX_ID)) %>%
  distinct(JustId) %>%
  left_join(datalink_min, by = "JustId") %>%
  left_join(riskstrat_min, by = "WlregAuditId") %>%
  summarise(
    n_unlinked_justids          = n(),
    n_absent_from_datalink      = sum(is.na(WlregAuditId)),
    n_in_datalink_no_wlreg_pxid = sum(!is.na(WlregAuditId) & is.na(px_id)),
    n_should_have_linked        = sum(!is.na(WlregAuditId) & !is.na(px_id))
  ) %>% print()

cat("\nINTERPRETATION:\n")
cat(" - n_forms_missing_PX_ID ~ 0  => bucket 1 is genuinely form-less; nothing missed.\n")
cat(" - large + n_absent_from_datalink large => JustId never in DataLink (upstream extract gap).\n")
cat(" - large + n_in_datalink_no_wlreg_pxid large => WlregAuditId -> px_id is the leak.\n")
cat(" - n_should_have_linked > 0 => justid_px_lookup_clean is dropping resolvable forms (bug).\n")
```

```{r}
# Rebuild the active patient-months and the matched set, then isolate bucket 1.
active_pm <- active_status_patient_months %>%
  distinct(PX_ID, MonthSnapshot, status_form)

matched_pm <- justform_patient_month %>%
  distinct(PX_ID, MonthSnapshot, status_form) %>%
  mutate(matched = TRUE)

# every linked StatN form for a patient (any status, any date)
all_linked_forms <- justform_status_long %>%
  filter(!is.na(PX_ID), !is.na(ChangeDate)) %>%
  select(PX_ID, JustId, ChangeDate, form_status = status_form)

# bucket 1 = active patient-month, no linked StatN form of ANY status for that PX_ID at all
bucket1 <- active_pm %>%
  left_join(matched_pm, by = c("PX_ID", "MonthSnapshot", "status_form")) %>%
  filter(is.na(matched)) %>%
  anti_join(all_linked_forms %>% distinct(PX_ID), by = "PX_ID") %>%   # truly zero StatN forms
  select(PX_ID, MonthSnapshot, listed_status = status_form)

cat("bucket-1 patient-months:", nrow(bucket1),
    " | distinct patients:", n_distinct(bucket1$PX_ID), "\n")

# distinct patient list (this is what you check by hand)
bucket1_patients <- bucket1 %>%
  distinct(PX_ID, listed_status) %>%
  arrange(PX_ID)

# For each bucket-1 patient, does a JustId exist for them ANYWHERE upstream?
#  - in JustFormHR (parent form table), via the JustId->PX_ID lookup
#  - in any raw StatN file, via that same lookup
justids_for_pxid <- justid_px_lookup_clean %>%
  semi_join(bucket1_patients, by = "PX_ID")

# JustIds that actually appear in the parent JustFormHR
justformhr_ids <- JustFormHR %>% mutate(JustId = as.character(JustId)) %>% distinct(JustId)
# JustIds that appear in any raw StatN file
statn_ids <- bind_rows(
  JustFormHRStat1 %>% transmute(JustId),
  JustFormHRStat2 %>% transmute(JustId),
  JustFormHRStat3 %>% transmute(JustId),
  JustFormHRStat4 %>% transmute(JustId)
) %>% mutate(JustId = as.character(JustId)) %>% distinct(JustId)

bucket1_audit <- bucket1_patients %>%
  left_join(
    justids_for_pxid %>%
      mutate(
        in_justformhr = JustId %in% justformhr_ids$JustId,
        in_any_statn  = JustId %in% statn_ids$JustId
      ) %>%
      group_by(PX_ID) %>%
      summarise(
        n_justids_linked   = n(),
        n_in_justformhr    = sum(in_justformhr),
        n_in_any_statn     = sum(in_any_statn),
        .groups = "drop"
      ),
    by = "PX_ID"
  ) %>%
  mutate(across(c(n_justids_linked, n_in_justformhr, n_in_any_statn),
                ~ tidyr::replace_na(., 0L)))

cat("\n=== BUCKET-1 PATIENT AUDIT (per patient) ===\n")
cat("n_justids_linked = how many JustIds map to this PX_ID via justid_px_lookup_clean\n")
cat("n_in_justformhr  = how many of those appear in the parent JustFormHR table\n")
cat("n_in_any_statn   = how many appear in any raw StatN criteria file\n\n")

# summary: how many bucket-1 patients have ZERO justids at all vs some
bucket1_audit %>%
  summarise(
    n_patients                  = n(),
    n_with_zero_justids         = sum(n_justids_linked == 0),
    n_with_justid_but_no_statn  = sum(n_justids_linked > 0 & n_in_any_statn == 0),
    n_with_statn_form           = sum(n_in_any_statn > 0)   # should be 0 by construction
  ) %>% print()

# the actual inspectable table
print(head(bucket1_audit, 50))

# write the full list to CSV so you can open it and spot-check against the SAS files
write.csv(bucket1_audit, "bucket1_patients_audit.csv", row.names = FALSE)
cat("\nFull list written to bucket1_patients_audit.csv\n")
```

```{r}
active_pm <- active_status_patient_months %>%
  distinct(PX_ID, MonthSnapshot, status_form)

matched_pm <- justform_patient_month %>%
  distinct(PX_ID, MonthSnapshot, status_form) %>%
  mutate(matched = TRUE)

# unmatched active patient-months, labeled by listed status
unmatched_by_status <- active_pm %>%
  left_join(matched_pm, by = c("PX_ID", "MonthSnapshot", "status_form")) %>%
  filter(is.na(matched)) %>%
  count(listed_status = status_form, name = "n_unmatched_pm") %>%
  left_join(
    active_pm %>% count(listed_status = status_form, name = "n_active_pm"),
    by = "listed_status"
  ) %>%
  mutate(pct_unmatched = round(100 * n_unmatched_pm / n_active_pm, 1)) %>%
  arrange(listed_status)

cat("\n=== UNMATCHED PATIENT-MONTHS BY LISTED STATUS ===\n")
print(unmatched_by_status)

cat("\nOnly the 'Adult Status 1/2/3/4' rows matter for your within-status bars.\n")
cat("Status 5/6 unmatched is expected and irrelevant (not in any within-status denominator).\n")
```

```{r}
# Unmatched active patient-months, restricted to Status 1-4 (the only ones in your bars).
active_pm_14 <- active_status_patient_months %>%
  filter(status_form %in% c("Adult Status 1","Adult Status 2","Adult Status 3","Adult Status 4")) %>%
  distinct(PX_ID, MonthSnapshot, status_form)

matched_pm <- justform_patient_month %>%
  distinct(PX_ID, MonthSnapshot, status_form) %>%
  mutate(matched = TRUE)

unmatched_14 <- active_pm_14 %>%
  left_join(matched_pm, by = c("PX_ID","MonthSnapshot","status_form")) %>%
  filter(is.na(matched)) %>%
  select(PX_ID, MonthSnapshot, listed_status = status_form)

# every linked StatN form for these patients (any status, any date)
all_forms <- justform_status_long %>%
  filter(!is.na(PX_ID), !is.na(ChangeDate)) %>%
  select(PX_ID, JustId, ChangeDate, form_status = status_form)

unmatched_14_anatomy <- unmatched_14 %>%
  left_join(all_forms, by = "PX_ID", relationship = "many-to-many") %>%
  group_by(PX_ID, MonthSnapshot, listed_status) %>%
  summarise(
    has_any_form_ever             = any(!is.na(JustId)),
    has_any_form_by_month         = any(ChangeDate <= MonthSnapshot),
    has_same_status_form_ever     = any(form_status == listed_status),
    has_same_status_form_by_month = any(form_status == listed_status & ChangeDate <= MonthSnapshot),
    .groups = "drop"
  ) %>%
  mutate(
    bucket = case_when(
      !has_any_form_ever                      ~ "1. no form ever (form-less)",
      has_same_status_form_by_month           ~ "0. same-status form by month (should have matched - BUG)",
      !has_any_form_by_month                  ~ "4. timing: only forms dated AFTER this month",
      has_same_status_form_ever               ~ "3a. timing: same-status form exists but dated AFTER this month",
      TRUE                                    ~ "2. form-less: has forms but never in listed status"
    ),
    kind = case_when(
      grepl("^0", bucket) ~ "BUG",
      grepl("^1|^2", bucket) ~ "form-less (structural)",
      grepl("^3a|^4", bucket) ~ "timing"
    )
  )

cat("\n=== UNMATCHED STATUS 1-4: bucket breakdown ===\n")
unmatched_14_anatomy %>% count(listed_status, bucket) %>%
  arrange(listed_status, bucket) %>% print(n = 100)

cat("\n=== UNMATCHED STATUS 1-4: timing vs form-less, by status ===\n")
unmatched_14_anatomy %>%
  count(listed_status, kind) %>%
  tidyr::pivot_wider(names_from = kind, values_from = n, values_fill = 0) %>%
  print()

# Is the timing component concentrated at the window's left edge? Show unmatched-by-month,
# split timing vs form-less, for Status 1-4 pooled.
cat("\n=== UNMATCHED STATUS 1-4 BY MONTH (timing vs form-less) ===\n")
unmatched_14_anatomy %>%
  count(MonthSnapshot, kind) %>%
  tidyr::pivot_wider(names_from = kind, values_from = n, values_fill = 0) %>%
  arrange(MonthSnapshot) %>%
  print(n = 100)
```

```{r}
# SELF-CONTAINED TEST. Does NOT modify the pipeline. It rebuilds active Status 1-4
# patient-months WITH their stathist episode dates, then compares three rules for
# matching a justification form to a (patient, month):
#
#   Rule A (current): form ChangeDate <= MonthSnapshot (first-of-month)
#   Rule B (within-month): form ChangeDate <= end of MonthSnapshot's calendar month
#   Rule C (within-episode): form ChangeDate <= the END of the stathist status episode
#                            that covers this MonthSnapshot
#
# For each rule we report how many active Status 1-4 patient-months remain UNMATCHED,
# split into timing (a same-status form exists for the patient, just later) vs
# form-less (no same-status form ever). Rule that collapses 'timing' to ~0 is the fix.

library(lubridate)

# active Status 1-4 patient-months, but KEEP episode begin/end so Rule C is possible
active_pm_dates <- stathist_monthly %>%
  filter(CANHX_STAT_label %in% c("Adult Status 1","Adult Status 2","Adult Status 3","Adult Status 4"),
         !is.na(PX_ID)) %>%
  transmute(
    PX_ID,
    MonthSnapshot,
    status_form = CANHX_STAT_label,
    episode_begin = CANHX_BEGIN_DT,
    episode_end   = CANHX_END_DT_adjusted
  ) %>%
  # one row per (PX_ID, month, status): keep the episode that covers this month
  group_by(PX_ID, MonthSnapshot, status_form) %>%
  arrange(desc(episode_begin), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup()

# same-status forms for these patients
forms <- justform_status_long %>%
  filter(!is.na(PX_ID), !is.na(ChangeDate)) %>%
  select(PX_ID, JustId, ChangeDate, status_form)

# join all same-status forms to each patient-month, then test each rule
joined <- active_pm_dates %>%
  left_join(forms, by = c("PX_ID","status_form"), relationship = "many-to-many") %>%
  mutate(
    month_end = ceiling_date(MonthSnapshot, "month") - days(1),
    ruleA = !is.na(ChangeDate) & ChangeDate <= MonthSnapshot,
    ruleB = !is.na(ChangeDate) & ChangeDate <= month_end,
    ruleC = !is.na(ChangeDate) & ChangeDate <= episode_end
  )

# per (patient, month): did ANY same-status form satisfy each rule?
pm_match <- joined %>%
  group_by(PX_ID, MonthSnapshot, status_form) %>%
  summarise(
    has_same_status_form_ever = any(!is.na(ChangeDate)),
    matched_A = any(ruleA),
    matched_B = any(ruleB),
    matched_C = any(ruleC),
    .groups = "drop"
  )

summarise_rule <- function(df, matched_col, rule_name) {
  df %>%
    summarise(
      rule              = rule_name,
      n_pm              = n(),
      n_matched         = sum(.data[[matched_col]]),
      n_unmatched       = sum(!.data[[matched_col]]),
      # of the unmatched, how many are timing (form exists, just not by the rule's cutoff)
      n_unmatched_timing   = sum(!.data[[matched_col]] & has_same_status_form_ever),
      # vs form-less (no same-status form at all)
      n_unmatched_formless = sum(!.data[[matched_col]] & !has_same_status_form_ever)
    )
}

cat("\n=== MATCH-RULE COMPARISON (active Status 1-4 patient-months) ===\n")
bind_rows(
  summarise_rule(pm_match, "matched_A", "A: <= first-of-month (current)"),
  summarise_rule(pm_match, "matched_B", "B: <= end of that month"),
  summarise_rule(pm_match, "matched_C", "C: <= end of stathist episode")
) %>% print()

# also by status, for rule C (the most permissive), to confirm it clears timing everywhere
cat("\n=== RULE C UNMATCHED, BY STATUS ===\n")
pm_match %>%
  group_by(status_form) %>%
  summarise(
    n_pm                 = n(),
    n_unmatched_timing   = sum(!matched_C & has_same_status_form_ever),
    n_unmatched_formless = sum(!matched_C & !has_same_status_form_ever),
    .groups = "drop"
  ) %>% print()
```

```{r}
library(lubridate)

active_pm_dates <- stathist_monthly %>%
  filter(CANHX_STAT_label %in% c("Adult Status 1","Adult Status 2","Adult Status 3","Adult Status 4"),
         !is.na(PX_ID)) %>%
  transmute(PX_ID, MonthSnapshot, status_form = CANHX_STAT_label,
            episode_begin = CANHX_BEGIN_DT, episode_end = CANHX_END_DT_adjusted) %>%
  group_by(PX_ID, MonthSnapshot, status_form) %>%
  arrange(desc(episode_begin), .by_group = TRUE) %>% slice(1) %>% ungroup()

forms <- justform_status_long %>%
  filter(!is.na(PX_ID), !is.na(ChangeDate)) %>%
  select(PX_ID, JustId, ChangeDate, status_form)

# For each patient-month, look at ALL same-status forms and find where they sit
# relative to this month and this episode.
resid <- active_pm_dates %>%
  left_join(forms, by = c("PX_ID","status_form"), relationship = "many-to-many") %>%
  group_by(PX_ID, MonthSnapshot, status_form, episode_begin, episode_end) %>%
  summarise(
    has_form_ever          = any(!is.na(ChangeDate)),
    matched_C              = any(!is.na(ChangeDate) & ChangeDate <= episode_end),
    # nearest same-status form date, and where it falls
    earliest_form          = suppressWarnings(min(ChangeDate, na.rm = TRUE)),
    latest_form            = suppressWarnings(max(ChangeDate, na.rm = TRUE)),
    any_form_after_episode = any(ChangeDate > episode_end, na.rm = TRUE),
    any_form_before_begin  = any(ChangeDate < episode_begin, na.rm = TRUE),
    any_form_within_episode= any(ChangeDate >= episode_begin & ChangeDate <= episode_end, na.rm = TRUE),
    .groups = "drop"
  )

# the residual = unmatched under C but DOES have a same-status form somewhere (the 1,583)
ruleC_residual <- resid %>%
  filter(!matched_C, has_form_ever) %>%
  mutate(
    why = case_when(
      any_form_within_episode               ~ "0. form within episode but C said no (BUG - investigate)",
      any_form_after_episode & !any_form_before_begin ~ "1. only same-status form(s) dated AFTER episode end",
      any_form_before_begin                 ~ "2. same-status form(s) only BEFORE this episode began (earlier episode)",
      TRUE                                   ~ "3. other"
    )
  )

cat("\n=== RULE C RESIDUAL (unmatched-but-has-same-status-form) ===\n")
ruleC_residual %>% count(status_form, why) %>% arrange(status_form, why) %>% print(n = 100)

cat("\n=== same, by month (is it left/right edge concentrated?) ===\n")
ruleC_residual %>% count(MonthSnapshot) %>% arrange(MonthSnapshot) %>% print(n = 100)

cat("\n=== how far is the gap? days between MonthSnapshot and nearest same-status form ===\n")
ruleC_residual %>%
  mutate(gap_days = as.numeric(pmin(abs(MonthSnapshot - earliest_form),
                                    abs(MonthSnapshot - latest_form)))) %>%
  summarise(
    n = n(),
    median_gap_days = median(gap_days, na.rm = TRUE),
    p90_gap_days    = quantile(gap_days, 0.9, na.rm = TRUE),
    max_gap_days    = max(gap_days, na.rm = TRUE)
  ) %>% print()
```

